### 2021-11-15 doc/appendix-pcod-sdmTMB.Rmd

---
title: "Annotated example of index standardization with sdmTMB"
output:
  pdf_document:
    highlight: tango
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618,
  cache = TRUE,
  autodep = TRUE
)
```

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(dplyr)
library(sdmTMB)
theme_set(theme_minimal())
```

We will fit a geostatistical spatiotemporal model with sdmTMB for the purposes of index standardization. We will use a data set built into the package: trawl survey data for Pacific Cod in Queen Charlotte Sound. The density units are kg/km^2^. Here, X and Y are coordinates in UTM zone 9.

First, we construct a mesh to approximate the spatial process, which is used in the SPDE calculations.
The `make_mesh()` function accepts the data coordinates and the minimum allowed distance between vertices (`cutoff`) or approximate number of knots in the mesh (`n_knots`).
Custom built meshes with `INLA` (including meshes with barriers limiting covariance across polygons) can also be passed in. See examples included in `?make_mesh`.

For example, a mesh with a cutoff distance of 10 km can be made with

```{r mesh, echo = TRUE, cache=FALSE}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)
```

```{r plot-mesh, fig.asp=0.9, out.width="4in"}
plot(mesh)
```

Next we will fit a GLMM (generalized linear mixed effects model). For this example, we will include `0 + as.factor(year)` so that there is a factor predictor that represents the mean estimate for each time slice. Setting `silent = FALSE` would give us optimization details as the model fits but clutter the R Markdown output.

```{r fit-pcod, warning=FALSE, message=FALSE}
fit <- sdmTMB(density ~ 0 + as.factor(year), data = pcod,
  time = "year", mesh = mesh, family = tweedie(link = "log")
)
```

Print the model to inspect the fit:

```{r print-fit}
print(fit)
```

Extract the parameter estimates as a data frame:

```{r tidy}
tidy(fit, conf.int = TRUE)
tidy(fit, effects = "ran_pars", conf.int = TRUE)
```

We can predict on the original data locations:

```{r predict}
pred <- predict(fit)
```

We can inspect randomized quantile residuals:

```{r qqnorm, out.width="5in"}
pred$resid <- residuals(fit)
qqnorm(pred$resid)
qqline(pred$resid)
```

```{r plot-resids}
ggplot(pred, aes(X, Y, colour = resid)) +
    geom_point() + facet_wrap(~year) +
    coord_fixed() + scale_colour_gradient2()
```

Next, we want to predict on a fine-scale grid over the entire survey domain so we can sum up the expected biomass across the entire region each year. There is a grid built into the sdmTMB package for Queen Charlotte Sound (named `qcs_grid`). Our prediction grid also needs to have all the covariates that we used in the model above. In this case, that is a `year` column as well as `X` and `Y`.

```{r predict-newdata}
pred_qcs <- predict(fit, newdata = qcs_grid)
```

Next we will make a small function to help plot maps.

```{r plot-map}
plot_map <- function(dat, column) {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}
```

There are four kinds of predictions that we get out of the model. First we will show the predictions that incorporate all fixed effects and random effects:

```{r plot-all-effects}
plot_map(pred_qcs, "exp(est)") +
  scale_fill_viridis_c(trans = "log10") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

We can also look at just the fixed effects, here `year`:

```{r plot-fix-defects}
plot_map(pred_qcs, "exp(est_non_rf)") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")
```

We can look at the spatial random effects that represent consistent deviations in space through time that are not accounted for by our fixed effects. In other words, these deviations represent consistent biotic and abiotic factors that are affecting biomass density but are not accounted for in the model.

```{r plot-spatial-effects}
plot_map(pred_qcs, "omega_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()
```

Finally, we can look at the spatiotemporal random effects that represent deviations from the fixed effect predictions and the spatial random effect deviations. These represent biotic and abiotic factors causing spatial correlation that are changing through time and are not accounted for in the model.

```{r plot-spatiotemporal-effects}
plot_map(pred_qcs, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()
```

The fastest way to assess uncertainty on our spatial predictions is with simulation from the joint precision matrix.

```{r sims}
pred_sims <- predict(fit, newdata = qcs_grid, sims = 500)
dim(pred_sims)
```

This returns a matrix with a row for each row of the `newdata` data frame and a column for each replicate draw. We can use this matrix to calculate the standard deviation or coefficient of variation:

```{r sims-cv-plot}
qcs_grid$cv <- apply(pred_sims, 1, function(x) sd(exp(x)) / mean(exp(x)))
ggplot(qcs_grid, aes(X, Y, fill = cv)) +
    geom_raster() + facet_wrap(~year) +
    coord_fixed() + scale_fill_viridis_c(trans = "log10", option = "D")
```

We can calculate our standardized population index by predicting on our survey domain grid and setting `return_tmb_object = TRUE`. This returns the TMB objective function, which is required to pass the output to `get_index()`.

```{r index, out.width="5in"}
qcs_grid$area <- 4 # all 2 x 2km
pred2 <- predict(
  fit, newdata = qcs_grid, return_tmb_object = TRUE,
  area = qcs_grid$area
)
ind <- get_index(pred2)
ggplot(ind, aes(year, est/1000)) + geom_line() +
  geom_ribbon(aes(ymin = lwr/1000, ymax = upr/1000), alpha = 0.4) +
  ylab("Biomass (t)")
```

The above example used the default `bias_correct = FALSE` in `get_index()`. This is fine for experimentation or a quick example, but will be biased due to the nonlinear (`exp()`) transformation of the random effects. Setting `bias_correct = TRUE` implements the generic by his correction algorithm that should correct for this, but can take considerable time and memory.

As an alternative, we can use the output from the joint precision matrix simulation draws to calculate our index. This should approximate the output of `get_index(x, bias_correct = TRUE)`. However, we note that there may be situations where this approach may fail or be biased.

```{r index-sims, out.width="5in"}
ind <- get_index_sims(pred_sims, area = rep(4, nrow(pred_sims)), level = 0.95)
ggplot(ind, aes(year, est/1000)) + geom_line() +
  geom_ribbon(aes(ymin = lwr/1000, ymax = upr/1000), alpha = 0.4) +
  ylab("Biomass (t)")
```