### 2021-11-18 doc/appendix-pcod-inlabru.Rmd
Fix a few small things; run styler

---
title: "Annotated example of index standardization with inlabru"
output:
  pdf_document:
    highlight: tango
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618,
  cache = TRUE,
  autodep = TRUE
)
```

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(dplyr)
library(sdmTMB)
library(INLA)
library(inlabru)
theme_set(theme_minimal())
```

This model is an extension of the same vignette using sdmTMB.

```{r mesh, echo = TRUE, cache=FALSE}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)

spde <- inla.spde2.pcmatern(
  mesh = mesh$mesh,
  prior.range = c(10, 0.05), # P(range < 10) = 0.05
  prior.sigma = c(2, 0.05)
) # P(sigma > 2) = 0.05
```

```{r plot-mesh, fig.asp=0.9, out.width="4in"}
plot(mesh)
```

There are a couple ways to include the fixed effects with `inlabru`. Unfortunately, using factor variables in the standard equation syntax does not work. In the first case, we can create a design matrix and add year effects individually to the formula.

We can fit the model first with `sdmTMB`. Note that this is a model with only the fixed year effects and a single spatial field, with presence-absence as the response.

```{r}
sdmtmb_fit <- sdmTMB(present ~ -1 + as.factor(year),
  data = pcod,
  time = "year", spde = mesh,
  family = binomial(link = "logit"),
  spatiotemporal = "off",
  priors = sdmTMBpriors(
    matern_st = pc_matern(
      range_gt = 10, range_prob = 0.05,
      sigma_lt = 2, sigma_prob = 0.05
    )
  )
)
sdmtmb_coefs <- tidy(sdmtmb_fit, effects = "fixed")
```

Now we can set up the design matrix for `inlabru`

```{r fit-pcod, warning=FALSE, message=FALSE}
m <- lm(present ~ -1 + as.factor(year), data = pcod)
model_mat <- model.matrix(m)
colnames(model_mat) <- paste0("x", unique(pcod$year))
dat <- cbind(pcod, model_mat)
```

Next we set up the `components` object for the model. This includes a formula with fixed or random effects (here only fixed) and the design of the spatial random field. The name `spatrf` can be anything, but includes the coordinates, spde model, and how the year effects are shared.

```{r}
components <- present ~ -1 + x2003 + x2004 + x2005 + x2007 +
  x2009 + x2011 + x2013 + x2015 + x2017 +
  spatrf(main = coordinates, model = spde)
# spatrf(main = coordinates,model = spde, group=year, ngroup=ncol(model_mat),
#       control.group = list(model = "iid"))

fit <- bru(
  components = components,
  family = "binomial", data = dat
)
bru_coefs <- fit$summary.fixed[, c("mean", "sd")]
```

Comparing the fixed effects between `sdmTMB` and `inlabru` shows the estimates are correlated, but not perfectly so

```{r}
plot(sdmtmb_coefs$estimate, bru_coefs$mean)
abline(a = 0, b = 1)
```

The other way to set up the fixed coefficients with `inlabru` is to use the `fac` argument. Note the location of the parameters changes from to `summary.random$fac`

```{r}
components <- present ~ 0 +
  fac(main = year, model = "factor_full") +
  spatrf(main = coordinates, model = spde)

fit2 <- bru(
  components = components,
  family = "binomial", data = dat
)
bru_coefs2 <- fit2$summary.random$fac[, c("mean", "sd")]
```

These estimates are perfectly correlated with the estimates above,

```{r}
plot(bru_coefs$mean, bru_coefs2$mean)
```