### 2021-12-09 doc/isec_abstract.Rmd
Edits

---
title: "ISEC_abstract"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Analyzing geostatistical data (coordinate-referenced observations) is becoming increasingly common in ecology; these data often include a discrete temporal dimension, such as monthly or yearly samples. Computational tools to model spatial and spatiotemporal processes have advanced rapidly over the last decade; examples of R packages for analyzing geostatistical data include mgcv, INLA and inlabru, VAST, and spBayes. 
SOMETHING HERE...
In this talk, we present a new R package sdmTMB. The goals of sdmTMB are not to replace functionality in existing packages, but provide a user-friendly implementation of SPDE models---similar to the popular R package glmmTMB, but with the addition of spatial and spatiotemporal models with an SPDE approach.
We extend the generalized linear mixed models (GLMMs) familiar to ecologists to include the following optional features:

* spatial fields
* spatiotemporal fields that may be independent by year or modeled with random walks or autoregressive processes
* smooth terms for covariates, using the familiar `s()` notation in mgcv
* spatially varying coefficient models (SVCs)
* non-stationary covariates (coefficients modeled as random walks)

Estimation is performed in sdmTMB via maximum marginal likelihood, implemented using Template Model Builder (TMB).
The sdmTMB package also allows for models to be passed to Stan, allowing for efficient Bayesian estimation.
We provide a general overview of the package structure, and demonstrate features using several examples including Snowy Owl winter irruptions in North America and a dataset describing the distribution of Pacific Cod in British Columbia. 
We highlight advantages of using sdmTMB for these classes of models including (1) estimation in TMB is often faster than alternative estimation routines, (2) sdmTMB provides a user-friendly interface that is familiar to many R users, (3) sdmTMB allows for easy out-of-sample model validation, (4) non-stationary processes are easily constructed with a formula interface (time-varying coefficients, spatially varying coefficient models). 
For analysts fitting species distribution models (SDMs) or estimating population abundance trends, sdmTMB provides specific functions for performing these calculations.