### 2021-12-11 doc/appendix-pcod-inlabru.Rmd
start adding detail to inlabru example

---
title: "Appendix 4: example of index standardization of fishery-independent survey data with inlabru"
output:
  pdf_document:
    highlight: tango
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618,
  cache = TRUE,
  autodep = TRUE
)
```

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(dplyr)
library(sdmTMB)
library(INLA)
library(inlabru)
theme_set(theme_minimal())
```

This model will attempt to replicate the model fit in Appendix 1 as an example of how index standardization can also be achieved with inlabru.
We will start by building an identical mesh using INLA.
These same functions that are used internally in the sdmTMB `make_mesh()` function.

```{r mesh, echo = TRUE, cache=FALSE}
loc_xy <- as.matrix(pcod[, c("X", "Y"), drop = FALSE])
inla_mesh <- INLA::inla.mesh.create(loc_xy, refine = TRUE, cutoff = 10)

spde <- inla.spde2.pcmatern(
  mesh = mesh$mesh,
  prior.range = c(10, 0.05), # P(range < 10) = 0.05
  prior.sigma = c(2, 0.05) # P(sigma > 2) = 0.05
) 
```

```{r plot-mesh, fig.asp=0.9, out.width="4in"}
plot(spde$mesh)
```

```{r mesh-sdmTMB, echo = TRUE, cache=FALSE}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)
```

There are a couple ways to include the fixed effects with `inlabru`. Unfortunately, using factor variables in the standard equation syntax does not work. In the first case, we can create a design matrix and add year effects individually to the formula.

We can fit the model first with `sdmTMB`. Note that this is a model with only the fixed year effects and a single spatial field, with presence-absence as the response.

```{r}
sdmtmb_fit <- sdmTMB(present ~ -1 + as.factor(year),
  data = pcod,
  time = "year", mesh = mesh,
  family = binomial(link = "logit"),
  spatiotemporal = "off",
  priors = sdmTMBpriors(
    matern_st = pc_matern(
      range_gt = 10, range_prob = 0.05,
      sigma_lt = 2, sigma_prob = 0.05
    )
  )
)
sdmtmb_coefs <- tidy(sdmtmb_fit, effects = "fixed")
```

Now we can set up the design matrix for `inlabru`

```{r fit-pcod, warning=FALSE, message=FALSE}
m <- lm(present ~ -1 + as.factor(year), data = pcod)
model_mat <- model.matrix(m)
colnames(model_mat) <- paste0("x", unique(pcod$year))
dat <- cbind(pcod, model_mat)
```

Next we set up the `components` object for the model. This includes a formula with fixed or random effects (here only fixed) and the design of the spatial random field. The name `spatrf` can be anything, but includes the coordinates, spde model, and how the year effects are shared.

```{r}
components <- present ~ -1 + x2003 + x2004 + x2005 + x2007 +
  x2009 + x2011 + x2013 + x2015 + x2017 +
  spatrf(main = coordinates, model = spde)
# spatrf(main = coordinates,model = spde, group=year, ngroup=ncol(model_mat),
#       control.group = list(model = "iid"))

fit <- bru(
  components = components,
  family = "binomial", data = dat
)
bru_coefs <- fit$summary.fixed[, c("mean", "sd")]
```

Comparing the fixed effects between `sdmTMB` and `inlabru` shows the estimates are correlated, but not perfectly so

```{r}
plot(sdmtmb_coefs$estimate, bru_coefs$mean)
abline(a = 0, b = 1)
```

The other way to set up the fixed coefficients with `inlabru` is to use the `fac` argument. Note the location of the parameters changes from to `summary.random$fac`

```{r}
components <- present ~ 0 +
  fac(main = year, model = "factor_full") +
  spatrf(main = coordinates, model = spde)

fit2 <- bru(
  components = components,
  family = "binomial", data = dat
)
bru_coefs2 <- fit2$summary.random$fac[, c("mean", "sd")]
```

These estimates are perfectly correlated with the estimates above,

```{r}
plot(bru_coefs$mean, bru_coefs2$mean)
```