### 2021-12-22 doc/appendix-validation.Rmd
Edit app.

---
title: "Appendix 6: sdmTMB model validation and speed-testing methods"
output:
  bookdown::pdf_document2:
    toc: true
number_sections: false
fig_caption: true
highlight: "monochrome"
csl: mee.csl
bibliography: refs.bib
header-includes:
  \usepackage[left]{lineno}
  \usepackage{bm}
  \usepackage{amssymb}
---

```{r setup, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  autodep = TRUE,
  cache.comments = FALSE
)
```

# Overall description of sdmTMB validation

The sdmTMB package has a number of validation processes to ensure the code is working as expected. 
This includes unit testing using testthat [@testthat] embedded within the package. 
This unit testing validates a large number of model configurations against simulated data and comparable models from other packages. 
We highlight some model validation in particular in this appendix.

In addition to the validation discussed here, Appendix 5 contains a comparison with VAST and Appendix 4 contains a comparison with INLA/inlabru.

This appendix illustrates a comparison of penalized smoothers with mgcv and a comparison of IID random intercepts with glmmTMB. 
Additionally, it illustrates some simulation testing of a spatial random field model with the main observation families within the sdmTMB package.

## Penalized smoothers

Here, we compare a number of smoother configurations against mgcv [@wood2017a]. 
Throughout the section, we will turn off spatial random fields so the models should be comparable.

### `s(x)`

First, a simple thin plate regression spline with a single variable. 
We will use simulated data from the mgcv package.

```{r, echo=TRUE}
library(sdmTMB)
```

```{r, echo=TRUE}
set.seed(19203)
# examples from ?mgcv::gam.models
dat <- mgcv::gamSim(3, n = 800)
m_mgcv <- mgcv::gam(y ~ s(x2), data = dat)
p_mgcv <- predict(m_mgcv)
dat$X <- runif(nrow(dat)) # fake
dat$Y <- runif(nrow(dat)) # fake
spde <- make_mesh(dat, c("X", "Y"), cutoff = 0.1)
m <- sdmTMB(y ~ s(x2),
  data = dat,
  mesh = spde, spatial = "off", spatiotemporal = "off"
)
p <- predict(m, newdata = NULL)
plot(p$est, p_mgcv)
abline(a = 0, b = 1)
cor(p$est, p_mgcv)
```

We can see the predictions match nearly exactly. 
Furthermore, we can use the in-development function `sdmTMB::plot_smooth()` to compared the smoothers themselves. 
Note that the sdmTMB version is showing the marginal effect of a prediction including the intercept whereas the mgcv version is showing only the smooth component. 
Nonetheless, here the smoothers look very similar except for these scale on the y axis.

```{r, echo=TRUE, fig.asp=1}
par(mfrow = c(2, 1))
plot_smooth(m)
plot(m_mgcv)
```

### `s(x1, by = x2)`

Next, we compare a model with a continuous `by` argument. 
This lets the `x2` smoother change depending on the value of `x1`.

```{r, echo=TRUE}
set.seed(19203)
# examples from ?mgcv::gam.models
# continuous by example:
dat <- mgcv::gamSim(3, n = 800)
m_mgcv <- mgcv::gam(y ~ s(x2, by = x1), data = dat)
p_mgcv <- predict(m_mgcv)
dat$X <- runif(nrow(dat)) # fake
dat$Y <- runif(nrow(dat)) # fake
spde <- make_mesh(dat, c("X", "Y"), cutoff = 0.1)
m <- sdmTMB(y ~ s(x2, by = x1),
  data = dat,
  mesh = spde, spatial = "off", spatiotemporal = "off"
)
p <- predict(m, newdata = NULL)
plot(p$est, p_mgcv)
abline(a = 0, b = 1)
cor(p$est, p_mgcv)
```

Again, the results are nearly identical.

### `s(x, y)`

Next, we compare a bivariate smoother, with near identical predictions.

```{r, echo=TRUE}
m_mgcv <- mgcv::gam(y ~ s(x2, x1), data = dat)
p_mgcv <- predict(m_mgcv)
m <- sdmTMB(y ~ s(x2, x1),
  data = dat,
  mesh = spde, spatial = "off", spatiotemporal = "off"
)
p <- predict(m, newdata = NULL)
plot(p$est, p_mgcv)
abline(a = 0, b = 1)
cor(p$est, p_mgcv)
```

## IID random intercepts

Here we compare a model that includes two IID random intercepts against the same model fit via glmmTMB [@brooks2017]. 
Again, we will turn off random fields so the models are comparable.

First, simulate some data. 
The variables `g` and `h` will represent groups g and h. 
g will have 50 groups. 
h will have 20 groups.

```{r, echo=TRUE}
predictor_dat <- data.frame(X = runif(1000), Y = runif(1000))
mesh <- make_mesh(predictor_dat, xy_cols = c("X", "Y"), cutoff = 0.1)
sim_dat <- sdmTMB_simulate(
  formula = ~ 1,
  data = predictor_dat,
  mesh = mesh,
  family = gaussian(),
  range = 0.5,
  phi = 0.1,
  sigma_O = 0.2,
  seed = 42,
  B = c(0.5) # B0 = intercept, B1 = a1 slope
)
sim_dat$g <- gl(50, 20)
iid_re_vals_g <- rnorm(50, 0, 0.3)
sim_dat$h <- gl(20, 50)
iid_re_vals_h <- rnorm(20, 0, 0.3)
sim_dat$observed <- sim_dat$observed + iid_re_vals_g[sim_dat$g] + iid_re_vals_h[sim_dat$h]
```

```{r, echo=TRUE}
m <- sdmTMB(observed ~ 1 + (1 | g) + (1 | h), mesh = mesh, data = sim_dat, 
  spatial = "off")

m_glmmTMB <- glmmTMB::glmmTMB(observed ~ 1 + (1 | g) + (1 | h), data = sim_dat)
```

We can extract the random effect estimates and compare them. 
We find them to be nearly identical:

```{r, echo=TRUE}
sdmTMB_pars <- tidy(m, "ran_pars")
glmmTMB_varcorr <- glmmTMB::VarCorr(m_glmmTMB)

sdmTMB_pars$estimate[sdmTMB_pars$term == "tau_G"][1]
sqrt(as.numeric(glmmTMB_varcorr$cond$g))

sdmTMB_pars$estimate[sdmTMB_pars$term == "tau_G"][2]
sqrt(as.numeric(glmmTMB_varcorr$cond$h))

sdmTMB_re <- as.list(m$sd_report, "Estimate")
glmmTMB_re <- glmmTMB::ranef(m_glmmTMB)$cond
plot(c(glmmTMB_re$g$`(Intercept)`, glmmTMB_re$h$`(Intercept)`), sdmTMB_re$RE,
  col = c(rep(1, 50), rep(2, 20)))
abline(0, 1)
```

The last plot illustrates the IID random effect deviations themselves with glmmTMB on the x-axis and sdmTMB on the y-axis. 
The 'g' deviations are black and the 'h' ones are red.

## Simulation-testing with various observation families

```{r}
knitr::include_graphics(here::here("figs", "sim-test-families-pars.pdf"))
```

```{r}
knitr::include_graphics(here::here("figs", "sim-test-families-b1.pdf"))
```

```{r}
knitr::include_graphics(here::here("figs", "sim-test-families-phi.pdf"))
```

```{r}
knitr::include_graphics(here::here("figs", "sim-test-families-range.pdf"))
```

```{r}
knitr::include_graphics(here::here("figs", "sim-test-families-sigmaO.pdf"))
```

# Supplemental methods for speed-testing related packages

TODO: Fig. 2 methods description