### 2022-01-06 doc/appendix-binomial-inla.Rmd
Tweak appendi

---
title: "Appendix 4: Dynamic species distribution model comparison between INLA, inlabru, and sdmTMB"
output:
  pdf_document:
    highlight: tango
    toc: true
    number_sections: true
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618,
  cache = TRUE,
  autodep = TRUE
)
```

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(dplyr)
library(sdmTMB)
library(INLA)
library(inlabru)
theme_set(theme_light())
```

In this example, we will fit a presence-absence species distribution model with fixed-effect intercepts for year and spatiotemporal random fields structured as AR(1).
We will use a binomial distribution and a logit link.
We will use a data set built into the package: trawl survey data for Pacific cod in Queen Charlotte Sound as described in earlier appendices.
INLA requires sequentially numbered time steps, we start by selecting the years we will include in this model and creating the year index variable `i_year`:

```{r}
d <- pcod %>% filter(year >= 2007)
yr_lu <- data.frame(year = unique(d$year), i_year = seq_along(unique(d$year)))
d <- dplyr::left_join(d, yr_lu)
```

In this example, we will use Penalized Complexity (PC) priors on the range and random field standard deviation ("sigma") for models built in INLA and sdmTMB.
The values amount to the probability the range is below a lower limit (e.g., `P(range < 5) = 0.05`) and that the sigma is above an upper limit (e.g., `P(sigma > 5) = 0.05`).

```{r set-priors, echo = TRUE, cache=FALSE}
range_min <- 5
sigma_max <- 5
prior_prob <- 0.05
```

In INLA a PC-prior can be used for the temporal autoregressive parameter, i.e., a prior that defines the probability the autocorrelation parameter is greater than a given value (e.g., `P(cor > 0) = 0.9`), but for now, sdmTMB only takes a Gaussian prior (e.g., `priors = sdmTMBpriors(ar1_rho = normal(0, 1))`).

```{r}
h_spec <- list(rho = list(prior = "pccor1", param = c(0, 0.9)))
```

We will start by building identical meshes using both sdmTMB and INLA.
Here, X and Y are coordinates in UTM zone 9.
These same functions are used internally in the sdmTMB `make_mesh()` function.

```{r sdm-mesh, echo = TRUE, cache=FALSE}
sdmTMB_mesh <- make_mesh(d, xy_cols = c("X", "Y"), cutoff = 10)
```

```{r inla-mesh, echo = TRUE, cache=FALSE}
loc_xy <- as.matrix(d[, c("X", "Y"), drop = FALSE])
inla_mesh <- INLA::inla.mesh.create(loc_xy, refine = TRUE, cutoff = 10)
spde <- inla.spde2.pcmatern(
  mesh = inla_mesh, # an smdTMB mesh could be inserted here with `sdmTMB_mesh$mesh`
  prior.range = c(range_min, prior_prob),
  prior.sigma = c(sigma_max, prior_prob)
)
```

Here, the meshes should be the same:

```{r plot-inla-mesh, fig.asp=0.9, out.height="2.5in"}
plot(sdmTMB_mesh)
title(main = "sdmTMB mesh")
plot(spde$mesh, main = "", asp = 1)
title(main = "INLA mesh")
```

# INLA

Data preparation for INLA models includes a formula, as well as objects made using several different functions including: `inla.spde.make.index()`, `inla.spde.make.A()`, and `inla.stack()`.
We use `control.inla = list(int.strategy = "eb", strategy = "gaussian")` for comparability with TMB.
We use weak priors on the factor fixed effects (`mean = 0, prec = 1/(100*100)`), which amounts to a `Normal(0, 100^2)` prior.

```{r inla}
tictoc::tic("INLA") # start timing
nyear <- length(unique(d$year))
iset <- inla.spde.make.index("i", n.spde = spde$n.spde, n.group = nyear)
A <- inla.spde.make.A(mesh = inla_mesh, loc = loc_xy, group = d$i_year)
sdat <- inla.stack(
  data = list(y = d$present),
  A = list(A, 1),
  effects = list(iset, year_factor = as.factor(d$year)),
  tag = "stdata"
)

formula <- y ~ 0 + year_factor + f(i,
  model = spde, group = i.group,
  control.group = list(model = "ar1", hyper = h_spec)
)

m_inla <- inla(formula,
  data = inla.stack.data(sdat),
  control.predictor = list(compute = TRUE, A = inla.stack.A(sdat)),
  family = "binomial",
  control.family = list(link = "logit"),
  control.fixed = list(expand.factor.strategy = "inla", mean = 0, prec = 1 / (100 * 100)),
  control.inla = list(int.strategy = "eb", strategy = "gaussian"),
  control.compute = list(config = TRUE)
)
tictoc::toc() # stop timing

summary(m_inla)
inla_coefs <- m_inla$summary.fixed[, c("mean", "sd")]
```

# inlabru

There are a couple ways to include the fixed effects with `inlabru`. Unfortunately, using factor variables in the standard equation syntax does not work.
One option is to create a design matrix and add year effects individually to the formula.
The other way to set up the fixed coefficients with inlabru is to use the `factor_full` (or `factor_contrast`) model (seen below within `fac()`).
This option causes the location of the factor fixed-effect parameter estimates to end up in `summary.random$fac`.

```{r}
dat <- as.data.frame(d)
sp::coordinates(dat) <- c("X", "Y")
dat <- as(dat, "SpatialPointsDataFrame")
```

```{r}
tictoc::tic("bru")
components <- present ~ 0 +
  f(
    main = coordinates, model = spde, group = i_year,
    control.group = list(model = "ar1", hyper = h_spec)
  ) +
  fac(main = year, model = "factor_full")

m_bru <- bru(
  components = components,
  family = "binomial",
  data = dat,
  options =
    list(
      bru_verbose = TRUE,
      control.fixed = list(expand.factor.strategy = "inla", mean = 0, prec = 1 / (100 * 100)),
      control.family = list(link = "logit"),
      control.inla = list(int.strategy = "eb", strategy = "gaussian")
    )
)
tictoc::toc()
```

```{r}
summary(m_bru)

m_inla$summary.hyperpar[, c("mean", "sd")]
m_bru$summary.hyperpar[, c("mean", "sd")]

bru_coefs <- m_bru$summary.random$fac[, c("ID", "mean", "sd")]
m_inla$summary.fixed[,c("mean", "sd")]
bru_coefs
```

# sdmTMB

We will now fit a nearly identical model with `sdmTMB()` (the AR(1) correlation parameter prior is different).
We again use very weak priors on the factor fixed effects.

```{r sdmtmb, warning=FALSE, message=FALSE}
tictoc::tic("sdmTMB")
m <- sdmTMB(present ~ 0 + as.factor(year),
  data = d,
  mesh = sdmTMB_mesh,
  time = "year",
  spatial = "off",
  spatiotemporal = "ar1",
  family = binomial(link = "logit"),
  priors = sdmTMBpriors(
    matern_st = pc_matern(
      range_gt = range_min, range_prob = prior_prob,
      sigma_lt = sigma_max, sigma_prob = prior_prob
    ),
    b = normal(0, 100),
    ar1_rho = normal(0, 2)
  ),
  reml = TRUE # to match INLA "eb" `int.strategy`
)
tictoc::toc()
```

```{r sdmtmb-coefs}
print(m)
tidy(m, "ran_pars", conf.int = TRUE)
sdmTMB_coefs <- tidy(m, effects = "fixed", conf.int = TRUE)
```

# Comparison

Comparing the fixed effects between `sdmTMB()` and `inla()` shows the estimates are very similar:

```{r coef-plot, out.width="5.5in"}
inla_coefs <- m_inla$summary.fixed[, c("0.025quant", "mean", "0.975quant")] %>% 
  rename(estimate = mean, conf.low = `0.025quant`, conf.high = `0.975quant`) %>% 
  mutate(term = unique(sdmTMB_coefs$term), model = "INLA")

bru_coefs <- m_bru$summary.random$fac[, c("0.025quant", "mean", "0.975quant")] %>% 
  rename(estimate = mean, conf.low = `0.025quant`, conf.high = `0.975quant`) %>% 
  mutate(term = unique(sdmTMB_coefs$term), model = "inlabru")

dplyr::bind_rows(mutate(sdmTMB_coefs, model = "sdmTMB"), inla_coefs, bru_coefs) %>% 
  ggplot(aes(x = estimate, y = term, xmin = conf.low, xmax = conf.high, colour = model)) +
  geom_pointrange(position = position_dodge(width = 0.2)) +
  theme(axis.title.y = element_blank())
```

With sdmTMB, we can extract simulated draws from the joint precision matrix using the `spread_sims()` or `gather_sims()` functions:

```{r sdmTMB-sims}
set.seed(1)
post_sdmTMB <- gather_sims(m, n_sims = 2000)
post_sdmTMB <- post_sdmTMB %>% 
  mutate(variable = 
      factor(`.variable`,
        levels = unique(post_sdmTMB$`.variable`),
        labels = c(seq(2007, 2017, 2), "Range", "AR1 rho", "Random Field SD")
      )
  )
```

We can then compare the distribution of these sdmTMB simulations to the marginal posterior densities from INLA.
From INLA, we can extract the marginal posterior densities using `inla.smarginal()` and name them to match the variable names from the sdmTMB `gather_sims()` function.
Alternatively, we could use `inla.posterior.sample(result = m_inla, n = 1000)` to sample from the INLA model posterior.

```{r inla-densities}
x <- list()
x[[1]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2007) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2007")
x[[2]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2009) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2009")
x[[3]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2011) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2011")
x[[4]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2013) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2013")
x[[5]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2015) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2015")
x[[6]] <- inla.smarginal(m_inla$marginals.fixed$year_factor2017) %>%
  as_tibble() %>%
  mutate(variable = "as.factor.year.2017")
x[[7]] <- inla.smarginal(m_inla$marginals.hyperpar$`Range for i`) %>%
  as_tibble() %>%
  mutate(variable = "range")
x[[8]] <- inla.smarginal(m_inla$marginals.hyperpar$`Stdev for i`) %>%
  as_tibble() %>%
  mutate(variable = "sigma_E")
x[[9]] <- inla.smarginal(m_inla$marginals.hyperpar$`GroupRho for i`) %>%
  as_tibble() %>%
  mutate(variable = "ar1_rho")
post_inla1 <- dplyr::bind_rows(x)
post_inla2 <- post_inla1 %>% mutate(variable = factor(variable,
  levels = unique(post_inla1$variable),
  labels = c(
    "2007", "2009", "2011", "2013", "2015", "2017",
    "Range", "Random Field SD", "AR1 rho"
  )
))
```

```{r compare-plot, out.width="6in"}
ggplot(post_inla2) +
  geom_histogram(
    data = post_sdmTMB, aes(.value, after_stat(density)), bins = 30,
    fill = "blue", alpha = 0.2
  ) +
  geom_line(data = post_inla2, aes(x, y), inherit.aes = FALSE) +
  facet_wrap(~variable, scales = "free")
```

# Annual distribution maps

We can predict on a grid (`qcs_grid`) provided with the `pcod` data in the sdmTMB package to map occurrence probabilities in each year in space.

```{r sdm-predictions}
nd <- filter(qcs_grid, year >= 2007)
pred <- predict(m, newdata = nd)
```

```{r maps, cache=FALSE, out.width="6in"}
ggplot(pred) +
  geom_raster(aes(X, Y, fill = plogis(est))) +
  facet_wrap(~year) +
  scale_fill_viridis_c() +
  coord_fixed()
```

Similar predictions are more complicated to extract from INLA, but are easy to plot for the original sample locations.

```{r inla-pred, out.width="6in"}
idx <- inla.stack.index(sdat, "stdata")$data
d$inla_pred <- m_inla$summary.fitted.values[idx, "mean"]
ggplot(d) +
  geom_point(aes(X, Y, colour = inla_pred), size = 2) +
  facet_wrap(~year) +
  scale_colour_viridis_c() +
  coord_fixed()
```

Or, we can make the prediction with inlabru. To do that, we must recreate `i_year`, create a `SpatialPixelsDataFrame` class data frame, and specify in `formula` to include both the random field (`f`), the factors (`fac`), and use the inverse-link transformation `plogis()`:

```{r inlabru-pred, out.width="6in"}
nd_bru <- nd
nd_bru <- dplyr::left_join(nd_bru, yr_lu)
sp::coordinates(nd_bru) <- c("X", "Y")
nd_bru <- as(nd_bru, "SpatialPixelsDataFrame")
pred_bru <- predict(m_bru, data = nd_bru, formula = ~ plogis(f + fac))
ggplot(as.data.frame(pred_bru)) +
  geom_raster(aes(X, Y, fill = mean)) +
  facet_wrap(~year) +
  scale_fill_viridis_c() +
  coord_fixed()
```