### 2022-11-02 doc/paper-jss/pcod.Rmd

# Example: spatiotemporal modelling of Pacific cod

```{r setup-pcod, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618,
  cache = TRUE,
  autodep = TRUE,
  cache.comments = FALSE,
  echo = TRUE
)
```

```{r packages, message=FALSE, warning=FALSE, cache=FALSE}
library(ggplot2)
library(dplyr)
library(sdmTMB)
theme_set(theme_minimal())
```

We will fit a geostatistical spatiotemporal model with sdmTMB for the purposes of index standardization. We will use a data set built into the package: trawl survey data for Pacific Cod in Queen Charlotte Sound. The density units are kg/km^2^. Here, X and Y are UTM zone 9 coordinates in kilometres. If our spatial coordinates were not already in an equidistant projection, we would transform them [e.g., using `sf::st_transform()`, @pebesma2018] to ensure that distance remains constant throughout the study region or using the built-in `sdmTMB::add_utm_columns()` function.

First, we construct a mesh to approximate the spatial process, which is used in the SPDE calculations. The `make_mesh()` function accepts the data coordinates and the minimum allowed distance between vertices (`cutoff`) or approximate number of knots in the mesh (`n_knots`). Custom built meshes with \pkg{INLA} (including meshes with barriers limiting covariance across polygons) can also be passed in. See examples included in `?make_mesh`.

For example, a mesh with a cutoff distance of 10 km can be made:

```{r mesh, echo = TRUE, cache=TRUE}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)
```

```{r plot-mesh, cache=TRUE, fig.align='center', fig.asp=1, out.width="3in", echo=TRUE}
plot(mesh)
```

Next we will fit a GLMM. For this example, we will include `0 + as.factor(year)` so that there is a factor predictor that represents the mean estimate for each time slice. Setting `silent = FALSE` would gives us optimization details as the model fits.

An alternative would be to model catch as the response with an offset (`offset` argument) for log area swept.

```{r fit-pcod, warning=FALSE, message=FALSE, cache=TRUE, echo=TRUE, results='hide'}
fit <- sdmTMB(
  density ~ 0 + as.factor(year),
  data = pcod,
  time = "year", 
  mesh = mesh,
  family = tweedie(link = "log"),
  silent = FALSE
)
```

Print the model to inspect the fit:

```{r print-fit, echo=TRUE}
print(fit)
```

Extract the parameter estimates as a data frame:

```{r pcod-tidy, echo=TRUE}
tidy(fit, conf.int = TRUE)
tidy(fit, effects = "ran_pars", conf.int = TRUE)
```

We can predict on the original data locations:

```{r pcod-predict, cache=TRUE, echo=TRUE}
pred <- predict(fit)
```

We can inspect simulated residuals from the DHARMa package:

```{r sim-resids, out.width="4in", warning=FALSE, results='hide', cache=TRUE, echo=TRUE}
set.seed(123)
simulate(fit, nsim = 500) %>%
  dharma_residuals(fit)
```

Or perhaps better, use MCMC residuals:

```{r mcmc-resids, warning=FALSE, results='hide', cache=TRUE}
set.seed(123)
r <- residuals(fit,
  type = "mle-mcmc", 
  mcmc_iter = 200, mcmc_warmup = 100
)
```

```{r mcmc-resids-vis, out.width="4in"}
qqnorm(r)
qqline(r)
```

Next, we want to predict on a fine-scale grid over the entire survey domain so we can sum up the expected biomass across the entire region each year. There is a grid built into the sdmTMB package for Queen Charlotte Sound (named `qcs_grid`). Our prediction grid also needs to have all the covariates that we used in the model above. In this case, that is a `year` column as well as `X` and `Y`.

```{r predict-newdata}
pred_qcs <- predict(fit, newdata = qcs_grid)
```

Next we will make a small function to help plot maps.

```{r plot-map}
plot_map <- function(dat, column) {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}
```

There are four kinds of predictions that we get out of the model. First we will show the predictions that incorporate all fixed effects and random effects:

```{r plot-all-effects}
plot_map(pred_qcs, "exp(est)") +
  scale_fill_viridis_c(trans = "log10") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

We can also look at just the fixed effects, here `year`:

```{r plot-fix-defects}
plot_map(pred_qcs, "exp(est_non_rf)") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")
```

We can look at the spatial random effects that represent consistent deviations in space through time that are not accounted for by our fixed effects. In other words, these deviations represent consistent biotic and abiotic factors that are affecting biomass density but are not accounted for in the model.

```{r plot-spatial-effects}
plot_map(pred_qcs, "omega_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()
```

Finally, we can look at the spatiotemporal random effects that represent deviations from the fixed effect predictions and the spatial random effect deviations. These represent biotic and abiotic factors causing spatial correlation that are changing through time and are not accounted for in the model.

```{r plot-spatiotemporal-effects}
plot_map(pred_qcs, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()
```

The fastest way to assess uncertainty on our spatial predictions is with simulation from the joint precision matrix.

```{r pcod-sims, cache=TRUE}
pred_sims <- predict(fit, newdata = qcs_grid, nsim = 200)
dim(pred_sims)
```

This returns a matrix with a row for each row of the `newdata` data frame and a column for each replicate draw. We can use this matrix to calculate the standard deviation or coefficient of variation:

```{r sims-cv-plot, cache=TRUE}
qcs_grid$cv <- apply(pred_sims, 1, function(x) sd(exp(x)) / mean(exp(x)))
ggplot(qcs_grid, aes(X, Y, fill = cv)) +
  geom_raster() +
  facet_wrap(~year) +
  coord_fixed() +
  scale_fill_viridis_c(trans = "log10", option = "D")
```

We can calculate a standardized population index by predicting on our survey domain grid and setting `return_tmb_object = TRUE`. This returns the \pkg{TMB} objective function, which is required to pass the output to `get_index()`.

```{r pcod-app-index, fig.asp=0.45, out.width="5in", cache=TRUE}
qcs_grid$area <- 4 # all 2 x 2km
pred2 <- predict(
  fit,
  newdata = qcs_grid, 
  return_tmb_object = TRUE
)
ind <- get_index(pred2, area = qcs_grid$area, bias_correct = TRUE)

ggplot(ind, aes(year, est / 1000)) +
  geom_line() +
  geom_ribbon(aes(ymin = lwr / 1000, ymax = upr / 1000), alpha = 0.4) +
  ylab("Biomass (t)")
```

The above example uses `bias_correct = TRUE` in `get_index()`. The `bias_correct` argument is a feature of TMB for models with random effects, and more details can be found in @thorson2016bias. Leaving `bias_correct = FALSE` is fine for experimentation, but will be biased due to the nonlinear (`exp()`) transformation of the random effects. Setting `bias_correct = TRUE` implements a generic bias correction algorithm that should correct for this, but takes more time and memory than skipping the bias correction.