### 2022-11-03 doc/paper-jss/inla-comparison.Rmd
figure captions for inla appendix

\clearpage

# Comparison between INLA, inlabru, and sdmTMB {#app:inla}

In this example, we will fit a presence-absence dynamic species distribution model with fixed-effect intercepts for year and spatiotemporal random fields structured as AR(1).
We will use a binomial distribution and a logit link.

We will start by loading necessary packages:

```{r packages-inla-comparison, message=FALSE, warning=FALSE, cache=FALSE, echo=TRUE}
library("ggplot2")
library("dplyr")
library("sdmTMB")
library("INLA")
library("inlabru")
theme_set(theme_light())
```

We will limit \pkg{INLA} to using a single thread to match how \pkg{TMB} is set up here. Both could optionally be run with parallel processing:

```{r inla-threads, echo=TRUE}
INLA::inla.setOption(num.threads = "1:1")
```

We will use a data set built into the package: trawl survey data for Pacific cod in Queen Charlotte Sound as described in earlier appendices.
\pkg{INLA} requires sequentially numbered time steps, we start by selecting the years we will include in this model and creating the year index variable `i_year`:

```{r pcod-inla-comparison-data, echo=TRUE}
d <- pcod %>% filter(year >= 2007)
yr_lu <- data.frame(year = unique(d$year), i_year = seq_along(unique(d$year)))
d <- dplyr::left_join(d, yr_lu, by = "year")
```

We will use Penalized Complexity (PC) priors on the range and random field standard deviation ("sigma") for models built in \pkg{INLA}, \pkg{inlabru}, and \pkg{sdmTMB}.
The values amount to the probability the range is below a lower limit (e.g., `P(range < 5) = 0.05`) and that the sigma is above an upper limit (e.g., `P(sigma > 5) = 0.05`).

```{r set-priors, echo = TRUE}
range_min <- 5
sigma_max <- 5
prior_prob <- 0.05
```

In INLA, a PC-prior can be used for the temporal autoregressive parameter, i.e., a prior that defines the probability the autocorrelation parameter is greater than a given value (e.g., `P(cor > 0) = 0.9`), but for now, sdmTMB only takes a Gaussian prior (e.g., `priors = sdmTMBpriors(ar1_rho = normal(0, 1))`).

```{r, echo=TRUE}
h_spec <- list(rho = list(prior = "pccor1", param = c(0, 0.9)))
```

We will start by building identical meshes using both sdmTMB and INLA.
Here, X and Y are coordinates in UTM zone 9.
These same functions are used internally in the sdmTMB `make_mesh()` function.

```{r sdm-mesh, echo = TRUE}
sdmTMB_mesh <- make_mesh(d, xy_cols = c("X", "Y"), cutoff = 10)
```

```{r inla-mesh, echo = TRUE}
loc_xy <- as.matrix(d[, c("X", "Y"), drop = FALSE])
inla_mesh <- INLA::inla.mesh.create(loc_xy, refine = TRUE, cutoff = 10)
spde <- inla.spde2.pcmatern(
  mesh = inla_mesh, # an smdTMB mesh could be inserted here with `sdmTMB_mesh$mesh`
  prior.range = c(range_min, prior_prob),
  prior.sigma = c(sigma_max, prior_prob)
)
```

Here, the meshes should be the same.

```{r plot-inla-mesh, fig.asp=0.9, out.height="2.5in",echo=FALSE, eval=FALSE}
plot(sdmTMB_mesh)
title(main = "sdmTMB mesh")
plot(spde$mesh, main = "", asp = 1)
title(main = "INLA mesh")
```

## INLA version

Data preparation for \pkg{INLA} models includes a formula, as well as objects made using several different functions including: `inla.spde.make.index()`, `inla.spde.make.A()`, and `inla.stack()`.
We use `control.inla = list(int.strategy = "eb", strategy = "gaussian")` for comparability with TMB.
We use weak priors on the factor fixed effects (`mean = 0, prec = 1/(100*100)`), which amounts to a `Normal(0, 100^2)` prior.

```{r inla, echo=TRUE}
tictoc::tic("INLA") # start timing
nyear <- length(unique(d$year))
iset <- inla.spde.make.index("i", n.spde = spde$n.spde, n.group = nyear)
A <- inla.spde.make.A(mesh = inla_mesh, loc = loc_xy, group = d$i_year)
sdat <- inla.stack(
  data = list(y = d$present),
  A = list(A, 1),
  effects = list(iset, year_factor = as.factor(d$year)),
  tag = "stdata"
)

formula <- y ~ 0 + year_factor + f(i,
  model = spde, group = i.group,
  control.group = list(model = "ar1", hyper = h_spec)
)

m_inla <- inla(formula,
  data = inla.stack.data(sdat),
  control.predictor = list(compute = TRUE, A = inla.stack.A(sdat)),
  family = "binomial",
  control.family = list(link = "logit"),
  control.fixed = list(
    expand.factor.strategy = "inla",
    mean = 0, prec = 1 / (100 * 100)
  ),
  control.inla = list(int.strategy = "eb", strategy = "gaussian")
)
tictoc::toc() # stop timing
```

We could use `summary()` to inspect the model, but will omit the output to save space.

```{r, echo=FALSE, eval=FALSE}
summary(m_inla)
```

## inlabru version

There are a couple ways to include the factor fixed effects with `inlabru`---using factor variables in the standard equation syntax does not work.
One option is to create a design matrix and add year effects individually to the formula.
The other way to set up the factor fixed coefficients with inlabru is to use the `factor_full` (or `factor_contrast`) model (seen below within `fac()`).
This option causes the location of the factor fixed-effect parameter estimates to end up in `summary.random$fac`.

```{r, echo=TRUE}
dat <- as.data.frame(d)
sp::coordinates(dat) <- c("X", "Y")
dat <- as(dat, "SpatialPointsDataFrame")
```

```{r, echo=TRUE}
tictoc::tic("bru")
components <- present ~ 0 +
  f(
    main = coordinates, model = spde, group = i_year,
    control.group = list(model = "ar1", hyper = h_spec)
  ) +
  fac(main = year, model = "factor_full")

m_bru <- bru(
  components = components,
  family = "binomial",
  data = dat,
  options =
    list(
      bru_verbose = TRUE,
      control.fixed = list(
        expand.factor.strategy = "inla",
        mean = 0, prec = 1 / (100 * 100)
      ),
      control.family = list(link = "logit"),
      control.inla = list(int.strategy = "eb", strategy = "gaussian")
    )
)
tictoc::toc()
```

We will again suppress the output of `summary()` to save space.

```{r, echo=TRUE, eval=FALSE}
summary(m_bru)
```

```{r, echo=TRUE}
m_bru$summary.random$fac[, c("ID", "mean", "sd", "0.025quant", "0.975quant")]
```

## sdmTMB version

We will now fit a nearly identical model with `sdmTMB()` (the AR(1) correlation parameter prior is different).
We again use very weak priors on the factor fixed effects.

```{r sdmtmb, warning=FALSE, message=FALSE, echo=TRUE}
tictoc::tic("sdmTMB")
m <- sdmTMB(present ~ 0 + as.factor(year),
  data = d,
  mesh = sdmTMB_mesh,
  time = "year",
  spatial = "off",
  spatiotemporal = "ar1",
  family = binomial(link = "logit"),
  priors = sdmTMBpriors(
    matern_st = pc_matern(
      range_gt = range_min, range_prob = prior_prob,
      sigma_lt = sigma_max, sigma_prob = prior_prob
    ),
    b = normal(0, 100),
    ar1_rho = normal(0, 2)
  ),
  reml = TRUE # to match INLA "eb" `int.strategy`
)
tictoc::toc()
```

```{r sdmtmb-m, echo=TRUE}
print(m)
```

We can most easily extract the fixed-effect estimates and confidence intervals using \pkg{smdTMB}'s `tidy()` function.

```{r sdmtmb-coefs, echo=TRUE}
sdmTMB_coefs <- tidy(m, effects = "fixed", conf.int = TRUE)
sdmTMB_coefs
```

## Comparison

The estimates for the spatial parameters are very similar between the INLA-based models and `sdmTMB()`:

```{r comparison, echo=TRUE}
# INLA
m_inla$summary.hyperpar[, c("mean", "sd", "0.025quant", "0.975quant")]

# inlabru
m_bru$summary.hyperpar[, c("mean", "sd", "0.025quant", "0.975quant")]

# sdmTMB
tidy(m, "ran_pars", conf.int = TRUE)
```

Note that, within \pkg{sdmTMB}, the standard errors for variance terms are not reported in the `tidy()` function as they could be misleading if considered in natural (rather than log) space.
The confidence intervals are, however, reported.

Comparing the fixed effects shows the estimates from all three models are also very similar (Fig.\ \ref{fig:coef-plot}).

```{r coef-plot, fig.asp=0.4, out.width="85%", echo=FALSE, fig.cap="Fixed effect estimates from all three models with 95 percent CI.", fig.pos='h', fig.align='center'}
inla_coefs <- m_inla$summary.fixed[, c("0.025quant", "mean", "0.975quant")] %>%
  rename(estimate = mean, conf.low = `0.025quant`, conf.high = `0.975quant`) %>%
  mutate(term = unique(sdmTMB_coefs$term), model = "INLA")

bru_coefs <- m_bru$summary.random$fac[, c("0.025quant", "mean", "0.975quant")] %>%
  rename(estimate = mean, conf.low = `0.025quant`, conf.high = `0.975quant`) %>%
  mutate(term = unique(sdmTMB_coefs$term), model = "inlabru")

dplyr::bind_rows(mutate(sdmTMB_coefs, model = "sdmTMB"), inla_coefs, bru_coefs) %>%
  ggplot(aes(x = estimate, y = term, xmin = conf.low, xmax = conf.high, colour = model)) +
  geom_pointrange(position = position_dodge(width = 0.2)) +
  theme(axis.title.y = element_blank())
```

With sdmTMB, we can extract simulated draws from the joint precision matrix using the `spread_sims()` or `gather_sims()` functions.

```{r sdmTMB-sims, echo=FALSE}
set.seed(1)
post_sdmTMB <- gather_sims(m, nsim = 2000)
post_sdmTMB <- post_sdmTMB %>%
  mutate(
    variable =
      factor(`.variable`,
        levels = unique(post_sdmTMB$`.variable`),
        labels = c(seq(2007, 2017, 2), "Range", "AR1 rho", "Random Field SD")
      )
  )
```

We can then compare the distribution of these \pkg{sdmTMB} simulations to the marginal posterior densities from \pkg{INLA}
From \pkg{INLA} (Fig.\ \ref{fig:compare-plot}). 
We do this by extracting the marginal posterior densities using `inla.smarginal()` and name them to match the variable names from the \pkg{sdmTMB} `gather_sims()` function.
Alternatively, we could use `inla.posterior.sample(result = m_inla, n = 1000)` to sample from the \pkg{INLA} model posterior assuming we had set\
`control.compute` ` = list(config = TRUE)`.

```{r inla-densities, echo=FALSE}
old_inla <- any(grepl(
  "year_factoryear_factor",
  names(m_inla$marginals.fixed)
))
yrs <- seq(2007, 2017, 2)
if (old_inla) {
  yf <- "year_factoryear_factor"
} else {
  yf <- "year_factor"
}
x <- list()
for (i in seq_along(yrs)) {
  x[[i]] <- inla.smarginal(m_inla$marginals.fixed[[paste0(yf, yrs[i])]]) %>%
    as_tibble() %>%
    mutate(variable = paste0("as.factor.year.", yrs[i]))
}
x[[7]] <- inla.smarginal(m_inla$marginals.hyperpar$`Range for i`) %>%
  as_tibble() %>%
  mutate(variable = "range")
x[[8]] <- inla.smarginal(m_inla$marginals.hyperpar$`Stdev for i`) %>%
  as_tibble() %>%
  mutate(variable = "sigma_E")
x[[9]] <- inla.smarginal(m_inla$marginals.hyperpar$`GroupRho for i`) %>%
  as_tibble() %>%
  mutate(variable = "ar1_rho")
post_inla1 <- dplyr::bind_rows(x)
post_inla2 <- post_inla1 %>% mutate(variable = factor(variable,
  levels = unique(post_inla1$variable),
  labels = c(yrs, "Range", "Random Field SD", "AR1 rho")
))
```

<!-- In this plot, the blue histograms are from \pkg{sdmTMB} and the density lines are from \pkg{INLA}: -->

```{r compare-plot, out.width="6in", fig.cap="Blue histograms represent simulated draws from the joint precision matrix of the sdmTMB model compared with the marginal posterior densities from the INLA model.", fig.align='center'}
ggplot(post_inla2) +
  geom_histogram(
    data = post_sdmTMB, aes(.value, after_stat(density)), bins = 30,
    fill = "blue", alpha = 0.2
  ) +
  geom_line(data = post_inla2, aes(x, y), inherit.aes = FALSE) +
  facet_wrap(~variable, scales = "free")
```

## Annual distribution maps

We can predict on a grid (`qcs_grid`) provided with the `pcod` data in the \pkg{sdmTMB} package to map occurrence probabilities in each year in space (Fig.\ \ref{fig:maps}).
For \pkg{sdmTMB}, the resulting values are in link space and require the appropriate inverse-link transformation (always contained in `model$family$linkinv()`, equivalent to inverse logit here), to produce occurrence probabilities:

```{r sdm-predictions}
nd <- filter(qcs_grid, year >= 2007)
pred <- predict(m, newdata = nd)
```

```{r maps, out.width="5.5in", fig.cap="Predicted occurence probabilities in time and space from the sdmTMB model.", fig.align='center'}
inverse_link <- m$family$linkinv
ggplot(pred) +
  geom_raster(aes(X, Y, fill = inverse_link(est))) +
  facet_wrap(~year) +
  scale_fill_viridis_c() +
  coord_fixed()
```

We can get similar predictions with \pkg{inlabru} using only a few additional steps: we must recreate `i_year`, create a `SpatialPixelsDataFrame` class data frame, and specify in `formula` to include both the random field (`f`), the factors (`fac`), and the appropriate inverse-link transformation (Fig.\ \ref{fig:inlabru-pred}).

```{r inlabru-pred, out.width="5.5in", fig.cap="Predicted occurence probabilities in time and space from the inlabru model.", fig.align='center'}
nd_bru <- nd
nd_bru <- dplyr::left_join(nd_bru, yr_lu, by = "year")
sp::coordinates(nd_bru) <- c("X", "Y")
nd_bru <- as(nd_bru, "SpatialPixelsDataFrame")
pred_bru <- predict(m_bru, data = nd_bru, formula = ~ plogis(f + fac))

ggplot(as.data.frame(pred_bru)) +
  geom_raster(aes(X, Y, fill = mean)) +
  facet_wrap(~year) +
  scale_fill_viridis_c() +
  coord_fixed()
```

Equivalent predictions are more complicated to extract from \pkg{INLA}, but are easy to plot for the original sample locations.

```{r inla-pred, out.width="5.5in"}
# idx <- inla.stack.index(sdat, "stdata")$data
# d$inla_pred <- m_inla$summary.fitted.values[idx, "mean"]
# 
# ggplot(d) +
#   geom_point(aes(X, Y, colour = inla_pred), size = 2) +
#   facet_wrap(~year) +
#   scale_colour_viridis_c() +
  # coord_fixed()
```