### 2023-08-04 doc/paper-jss/paper-jss.Rmd
In progress revamp of examples Requires very latest sdmTMB

---
documentclass: jss
author:
  - name: Sean C. Anderson
    orcid: 0000-0001-9563-1937
    affiliation: 'Pacific Biological Station'
    address: |
      | Fisheries and Oceans Canada, Nanaimo, BC, Canada
    email: \email{sean.anderson@dfo-mpo.gc.ca}
  - name: Eric J. Ward
    orcid: 0000-0002-4359-0296
    affiliation: 'Northwest Fisheries Science Center \AND'
    address: |
      | National Oceanic and Atmospheric Administration, National Marine Fisheries Service, Seattle, WA, USA
  - name: Philina A. English
    orcid: 0000-0003-2992-6782
    affiliation: 'Pacific Biological Station'
    address: |
      | Fisheries and Oceans Canada, Nanaimo, BC, Canada
  - name: Lewis A. K. Barnett
    orcid: 0000-0002-9381-8375
    affiliation: 'Alaska Fisheries Science Center'
    address: |
      | National Marine Fisheries Service, National Oceanic and Atmospheric Administration, Seattle, WA, USA
bibliography: refs.bib
title:
  formatted: "\\pkg{sdmTMB}: An \\proglang{R} Package for Fast, Flexible, and User-Friendly Generalized Linear Mixed Effects Models with Spatial and Spatiotemporal Random Fields"
  # If you use tex in the formatted title, also supply version without
  plain:     "sdmTMB: An R Package for Fast, Flexible, and Accessible Generalized Linear Mixed Effects Models with Spatial and Spatiotemporal Random Fields"
  # For running headers, if needed
  short:     "\\pkg{sdmTMB}: geostatistical SPDE-based GLMMs with \\pkg{TMB}"
abstract: >
  Geostatistical data are common across scientific fields and available in increasingly large datasets. However, appropriate models to analyse these data, such as generalised linear mixed effects models with Gaussian random fields, are computationally intensive and challenging to implement for many users. Here, we introduce the \proglang{R} package \pkg{sdmTMB}, which extends the flexible interface familiar to users of \pkg{lme4}, \pkg{glmmTMB}, or \pkg{mgcv} to include spatial and spatiotemporal latent fields using a predictive-process SPDE-(stochastic partial differential equation) based approach. SPDE matrices are constructed with \proglang{R}-\pkg{INLA} and estimation is conducted via maximum marginal likelihood with Template Model Builder (\pkg{TMB}) or via Bayesian inference with \pkg{tmbstan} and \pkg{rstan}. We describe the model and explore case studies that illustrate \pkg{sdmTMB}'s flexibility in implementing penalised smoothers, non-stationary processes (time-varying and spatially varying coefficients), and anisotropy (directionally dependent spatial correlation). Additional functionality includes hurdle models, break-point effects, and out-of-sample cross validation. Finally, we compare the functionality, speed, and interfaces of related software, demonstrating that \pkg{sdmTMB} can be an order of magnitude faster than \proglang{R}-\pkg{INLA}. We hope \pkg{sdmTMB}'s accessible interface will help open this useful class of models to a wider field of geostatistical analysts.
keywords:
  # at least one keyword must be supplied
  formatted: [Gaussian Markov random fields (GMRF), generalized linear mixed effects models (GLMM), INLA, SPDE, species distribution modelling, R package, spatio-temporal, spatial-temporal, Template Model Builder]
  plain:     [Gaussian Markov random fields (GMRF), generalized linear mixed effects models (GLMM), INLA, SPDE, species distribution modelling, R package, spatio-temporal, spatial-temporal, Template Model Builder]
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{bm}
  \usepackage{pdflscape}
  \usepackage{tablefootnote}
  \usepackage{threeparttable}
  \usepackage{booktabs}
  \usepackage{pifont}
  \usepackage{newunicodechar}
  \usepackage{gensymb}
  \newunicodechar{✓}{\ding{51}}
  \newunicodechar{✗}{\ding{55}}
  \newunicodechar{˚}{\degree}
  \DeclareGraphicsExtensions{.pdf,.png,.jpg}
  \widowpenalty10000
  \clubpenalty10000
output: rticles::jss_article
---

```{r main-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  # prompt = FALSE,
  cache = TRUE,
  autodep = TRUE,
  fig.width = 7,
  fig.asp = 0.618,
  fig.pos = "ht",
  cache.comments = TRUE,
  dev = "pdf",
  dpi = 140,
  optipng = "-strip all"
  # R.options = list(prompt = "R> ", continue = "+ ")
)
knitr::knit_hooks$set(optipng = knitr::hook_optipng)
options(prompt = "R> ", continue = "+  ", width = 72, useFancyQuotes = FALSE)
```

\clearpage
# Introduction

Data are often collected in space or in space repeatedly over time.
While such data are a rich source of information [e.g., @legendre1989a; @rossi1992; @tilman1997spatial], they are challenging to properly model---data closer in space and time are usually more similar to each other than data farther apart due to measured and unmeasured variables [@cressie1993; @diggle2007; @cressie2011].
While measured variables can be accounted for with predictors in a model (e.g., measuring and modelling temperature effects on species abundance), unmeasured variables (e.g., everything influencing species abundance but not explicitly modelled) can cause residual spatial correlation.
Accounting for this residual correlation is important because doing so allows for valid statistical inference [@legendre1989a; @dormann2007], can improve predictions  [e.g., @shelton2014], and can be of interest itself [e.g., @thorson2019d; @barnett2021].

Geostatistical generalized linear mixed effects models (GLMMs) with spatially correlated random effects are a class of models appropriate for these data [@rue2005gmrf; @diggle2007; @cressie2011].
Similarly to how random intercepts can account for correlation among groups, spatial or spatiotemporal random effects can account for unmeasured variables causing observations to be correlated in space or space and time.
A common approach to modelling these spatial effects is with Gaussian random fields (GRFs), where the random effects describing the spatial patterning are assumed to be drawn from a multivariate normal (MVN) distribution, constrained by some covariance function such as the exponential or Matérn [@cressie1993; @diggle2007; @chiles1999].

Such models quickly become computationally limiting due to the need to invert large matrices to keep track of covariation among data [e.g., @rue2005gmrf; @latimer2009].
Many solutions have been proposed, such as predictive processes [@banerjee2008; @latimer2009], the stochastic partial differential equation (SPDE) approximation to GRFs [@lindgren2011], and nearest-neighbour Gaussian processes [@datta2016; @finley2022].
These approaches reduce the scale of the covariance estimation problem while providing a means to evaluate the data likelihood, thereby allowing fitting via Bayesian [@gelfand2017] or maximum likelihood methods.
This can greatly improve computational efficiency [e.g., @heaton2019].
The SPDE approach is a solution popularized via the \proglang{R}-\pkg{INLA} package [@rue2009; @lindgren2011; @lindgren2015] and an implementation in \pkg{TMB} [Template Model Builder, @kristensen2016] that relies on \proglang{R}-\pkg{INLA} to create input matrices [e.g., @thorson2019; @osgood-zimmerman2021].
Details are beyond the scope of this paper and are not necessary to use the software discussed here, but the idea is that the solution to a specific SPDE is a GRF with a Matérn covariance function and this 'trick' enables one to efficiently fit approximations to GRFs to large spatial datasets [@lindgren2011].

Systems for specifying statistical models that can include the SPDE, such as \proglang{R}-\pkg{INLA} and \pkg{TMB}, are flexible and powerful but are challenging to use for many applied researchers.
For example, \pkg{TMB} requires the user to program in a \proglang{C++} template and it can be slow to experiment with multiple models when writing bespoke model code.
Packages such as \pkg{lme4} [@bates2015] and \pkg{glmmTMB} [@brooks2017] let users quickly iterate and explore statistical models---focusing on evaluating fit and comparing models---but do not have built-in SPDE functionality.
Packages such as \pkg{VAST} [@thorson2019] and \pkg{inlabru} [@bachl2019] are powerful user interfaces to fit spatial models that use the SPDE, but they either lack a modular interface familiar to those who have used \pkg{lme4} or \pkg{glmmTMB}, or lack some functionality.
We provide a more detailed comparison of related software packages in Table \ref{tab:functionality} and the Discussion.

Here, we introduce the \proglang{R} package \pkg{sdmTMB}, which implements geostatistical spatial and spatiotemporal GLMMs using \pkg{TMB} for model fitting and \proglang{R}-\pkg{INLA} to set up SPDE matrices.
Our aim is not to replace the above-mentioned statistical packages, but to provide a fast, flexible, and user-friendly interface that is familiar to users of \pkg{lme4}, \pkg{glmmTMB}, or \pkg{mgcv} [@wood2017a] for a specific class of spatial and spatiotemporal models.
One common application in the field of ecology is species distribution models (SDMs), hence the package name (i.e., SDMs with \pkg{TMB}: \pkg{sdmTMB}).
This paper describes the basic functionality of this \proglang{R} package and its underlying statistical model, illustrates its use through two case studies, and concludes with a comparison to related software.

# Model description

\pkg{sdmTMB} fits GLMMs to spatial or spatiotemporal geostatistical data.
Geostatistical data refers to data observed at specific spatial coordinates reflecting some underlying spatial process [@rossi1992; @diggle2007].
These data can be collected across discrete points in time.
Areal data (data aggregated to polygon or grid level) may be analyzed using other spatial models, including conditional (CAR) autoregressive models [e.g., @verhoef2018];
\pkg{sdmTMB} can also fit models with areal data if each polygon has an associated centroid.
A benefit of the geostatistical approach over CAR or similar models is that the parameters describing spatial covariance can be more easily interpreted [@wall2004] (e.g., spatial variance and range---the distance at which points are effectively independent).

The process component of an \pkg{sdmTMB} model can be formed by any combination of main ("fixed") effects, spatial intercept random fields, spatiotemporal intercept random fields, IID (independent and identically distributed) random intercepts (but not currently random slopes), time-varying effects, and spatially varying effects (Fig.\ \ref{fig:diagram}).
This process component is combined with an observation error family (e.g., Gaussian, Gamma, binomial, Tweedie) and link (e.g., identity, log, logit) as in any generalized linear model (GLM) [@mccullagh1989].
Some families can be combined into a two-part "delta" or "hurdle" model [@aitchison1955] to model the zero vs. non-zero observations separately from the positive observations.
See the model description vignette for a full description of the statistical model (`browseVignettes("sdmTMB")` or see the \href{https://pbs-assess.github.io/sdmTMB/articles/}{rendered version}).

The GLMMs underpinning \pkg{sdmTMB} models are spatially explicit---they estimate interpretable parameters of a spatial covariance function: parameters defining the magnitude of spatial variation and the rate of correlation decay with distance.
In contrast, semi- and non-parametric approaches do not estimate spatial covariance functions (e.g., \pkg{randomForest} [@liaw2002], \pkg{MaxEnt} [@phillips2006], and most smooths in \pkg{mgcv} [@wood2017a]).
The random fields in \pkg{sdmTMB} are structured as MVN constrained by a Matérn covariance function [@matern1960].
The Matérn can accommodate a range of shapes and can be both isotropic (correlation decays the same in all directions) or anisotropic (correlation is directionally dependent) [@haskard2007].
The Matérn standard deviations are estimated separately for the various fields and the range---the distance at which spatial correlation decays to $\sim 0.13$ [@lindgren2015]---can be shared or estimated separately (`share_range` argument).

By default, if spatiotemporal fields are included, they are assumed IID; however, additional options allow them to be modelled as a random walk or first-order autoregressive, AR(1), process (Fig.\ \ref{fig:diagram}).
Turning off both spatial and spatiotemporal effects allows for comparison with standard non-spatial GLMs or GLMMs.
We include additional flexibility in specifying the linear fixed effect matrix: covariates can be modelled as penalized smooth functions (generalized additive models, GAMs) using the same `s()` syntax as in \pkg{mgcv} [@wood2017a] (thereby allowing automatic selection of smoother 'wiggliness'), or can be modelled with threshold shapes: hockey-stick models, `breakpt()`, [@barrowman2000] or logistic functions, `logistic()`.
\pkg{sdmTMB} can use many `mgcv::s()` and `mgcv::t2()` smoothers including bivarate smoothers (`s(x, y)`), smoothers varying by continuous or categorical variables (`s(x1, by = x2)`), cyclical smoothers (`s(x, bs = "cc")`), and smoothers with specified basis dimensions (`s(x, k = 4)`) [@wood2017a].

The \pkg{sdmTMB} model is fit by maximum marginal likelihood.
Internally, a \pkg{TMB} [@kristensen2016] model template is used to calculate the marginal log likelihood and its gradient, and the negative log likelihood is minimized via the non-linear optimization routine `stats::nlminb()` in \proglang{R} [@gay1990; @r2021].
Random effects are estimated at values that maximize the log likelihood conditional on the estimated fixed effects and are integrated over via the Laplace approximation [@kristensen2016].
After rapid model exploration with maximum likelihood, one can optionally pass an \pkg{sdmTMB} model to the \proglang{R} package \pkg{rstan} [@carpenter2017; @rstan] via \pkg{tmbstan} [@monnahan2018] to estimate the joint posterior distribution for Bayesian inference.

\pkg{sdmTMB} models can include penalized likelihoods by assigning priors to model parameters (`?sdmTMBpriors`).
These priors may be useful in cases where estimation is difficult because of identifiability issues or relatively flat likelihood surfaces, or to impart prior information or achieve regularization.
Following other recent SPDE implementations in \pkg{TMB} [@breivik2021; @osgood-zimmerman2021], penalized complexity (PC) priors [@simpson2017; @fuglstad2019] (`?pc_matern`) can constrain the spatial range and variance parameters.

The results from an \pkg{sdmTMB} model can be used to generate various quantities of interest.
For example, for the field of ecology, \pkg{sdmTMB} provides functionality for generating population-level summaries for each time slice including the center of gravity  (`get_cog()`) [e.g., @thorson2016cog] and total population index (`get_index()`) given a user-supplied grid.

\begin{landscape}
\begin{figure}[p]
\centering
\includegraphics[width=9in]{../../figs/diagram}
\caption[]{
Components of an \pkg{sdmTMB} model with illustrations, descriptions, examples, notation, and example code.
An \pkg{sdmTMB} model can be built from any combination of the process components (first six rows) plus an observation component (last row).
The examples are from an SDM context, but the model can be fit to any spatially referenced point data.
Notation:
We refer to design matrices as $\bm{X}$.
The indexes $\bm{s}$, $t$, and $g$ index spatial coordinates, time, and group, respectively.
The $\sigma$ and $\bm{\Sigma}$ symbols represent standard deviations and covariance matrices, respectively.
All other symbols refer to the  described model components (e.g., $\bm{\beta}$ and $\bm{\omega}$ refer to a vector of main effects and spatial random field deviations, respectively).
See the \pkg{sdmTMB} \href{https://pbs-assess.github.io/sdmTMB/articles/}{model description vignette} for a full description of the model.
Note that \code{s()} denotes a smoother as in \pkg{mgcv} \citep{wood2017a}, \code{breakpt()} denotes a breakpoint 'hockey-stick' shape \citep[e.g.][]{barrowman2000}, \code{(1|g)} denotes a random intercept by group \code{g}, and \code{\textasciitilde\ 0} is used in an \proglang{R} formula to omit an intercept.}\label{fig:diagram}
\end{figure}
\end{landscape}

# Introductory example using fish survey data

## Illustrating the need for spatial random fields

We begin with a simple species distribution model of encounter probability of
Pacific Cod (TODO spp) from a trawl survey conducted off the west
coast of Vancouver Island, British Columbia, Canada. We will fit a logistic
regression of encounter probability with and without spatial random fields to
illustrate the importance of accounting for spatial correlation.

```{r libs, warning=FALSE, message=FALSE, echo=TRUE, cache=FALSE}
library(sdmTMB)
library(dplyr)
library(ggplot2)
```

```{r libs-extras}
theme_set(theme_light())
options(ggplot2.continuous.fill = "viridis")
```

```{r dog-binomial, results='hide', message=FALSE, warning=FALSE, echo=TRUE}
mesh_pcod <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 8)
plot(mesh_pcod)
```

```{r}
fit_bin_rf <- sdmTMB(
  present ~ poly(log(depth), 2),
  data = dat, mesh = mesh, spatial = "on",
  family = binomial(link = "logit")
)
fit_bin <- update(fit_bin_rf, spatial = "off")
```

```{r dog-bin-summary, echo=TRUE}
summary(fit_bin_rf)
```

Describe what we see here

We can test for spatial autocorrelation with a Moran's I test or with a visual inspection of the residuals.

```{r}
# set.seed(1)
# pcod$resid_bin <- residuals(fit_bin)
# set.seed(1)
# pcod$resid_bin_rf <- residuals(fit_bin_rf)
# g1 <- ggplot(pcod, aes(X, Y, colour = resid_bin)) + geom_point() +
#   scale_color_gradient2() +
#   theme(legend.position = "bottom")
# g2 <- ggplot(pcod, aes(X, Y, colour = resid_bin_rf)) + geom_point() +
#   scale_color_gradient2() +
#   theme(legend.position = "bottom")
# cowplot::plot_grid(g1, g2, ncol = 2)
```

```{r, echo=FALSE, results='hide'}
test_autocor <- function(obj) {
  set.seed(1)
  s <- simulate(obj, nsim = 500)
  pr <- predict(obj, type = "response")$est
  r <- DHARMa::createDHARMa(
    simulatedResponse = s,
    observedResponse = pcod$present,
    fittedPredictedResponse = pr
  )
  DHARMa::testSpatialAutocorrelation(r, x = pcod$X, y = pcod$Y, plot = FALSE)
}
test_autocor(fit_bin)
test_autocor(fit_bin_rf)
```

We can also see that the AIC is considerably lower:

```{r}
AIC(fit_bin_rf, fit_bin)
```

Further testing could be done with cross validation via `sdmTMB_cv()`.

```{r}
cv_bin_rf <- sdmTMB_cv(present ~ poly(log(depth), 2),
  data = dat, mesh = mesh, spatial = "on",
  family = binomial(link = "logit"), k_folds = 10
)
cv_bin_rf <- sdmTMB_cv(present ~ poly(log(depth), 2),
  data = dat, mesh = mesh, spatial = "off",
  family = binomial(link = "logit"), k_folds = 10
)
```

Indeed, the log likelihood of left-out data is considerably larger for the model that includes random fields:

```{r}
cv_bin_rf$sum_loglik
cv_bin_rf$sum_loglik
```

The standard errors on our fixed effects have increased:

```{r}
tidy(fit_bin_rf, conf.int = T)
tidy(fit_bin, conf.int = T)
```

We now have an effect of depth and a spatial random field, which combined in link space result in our overall prediction. Note the difference between only the fixed effects (left) and including the spatial random field (right).

```{r}
p <- predict(fit_bin_rf, newdata = qcs_grid)
plot_spatial_map <- function(dat, column, title) {
  ggplot(dat, aes(X, Y, fill = {{ column }})) +
    geom_raster() +
    coord_fixed() +
    theme(legend.position = "bottom") +
    ggtitle(title)
}
g1 <- plot_spatial_map(p, plogis(est_non_rf), "Fixed effects")
g2 <- plot_spatial_map(p, omega_s, "Spatial random field") +  scale_fill_gradient2()
g3 <- plot_spatial_map(p, plogis(est), "Combined prediction")
cowplot::plot_grid(g1, g2, g3, ncol = 3)
```

# Spatiotemporal species distribution modeling with dogfish

```{r dog-head-dat, echo=TRUE}
dat <- select(dogfish, longitude, latitude, year, catch_weight, area_swept, depth)
head(dat)
```

```{r dog-utms, echo=TRUE}
dat <- add_utm_columns(dat, c("longitude", "latitude"), units = "km", utm_crs = 32609)
dat$log_depth <- log(dat$depth)
```

```{r dog-mesh, echo=TRUE}
mesh <- make_mesh(dat, xy_cols = c("X", "Y"), cutoff = 8)
plot(mesh)
```

```{r dog-tw, results='hide', message=FALSE, warning=FALSE, echo=TRUE}
fit_tw <- sdmTMB(
  catch_weight ~ s(log_depth),
  data = dat,
  mesh = mesh,
  offset = log(dat$area_swept),
  time = "year",
  time_varying = ~ 1,
  time_varying_type = "ar1",
  spatial = "on",
  spatiotemporal = "iid",
  anisotropy = TRUE,
  share_range = TRUE,
  silent = FALSE,
  family = tweedie()
)
```

We can use `update()` to refit our model with four additional families:
delta-gamma, delta-lognormal, delta-Poisson-link-gamma, and
delta-Poisson-link-lognormal.

```{r dog-update, results='hide', message=FALSE, warning=FALSE, echo=TRUE}
fit_dg <- update(fit_tw, family = delta_gamma(), spatiotemporal = list("off", "iid"))
fit_dl <- update(fit_dg, family = delta_lognormal())
fit_dpg <- update(fit_dg, family = delta_poisson_link_gamma())
fit_dpl <- update(fit_dg, family = delta_poisson_link_lognormal())
```

We can then compare the fits via AIC:

```{r dog-aic, echo=TRUE}
AIC(fit_tw, fit_dg, fit_dl, fit_dpg, fit_dpl) |> 
  mutate(delta_AIC = AIC - min(AIC)) |> 
  arrange(delta_AIC)
```

We find that the Poisson-link-lognormal model is favoured with AIC. We next
test two additional model formulations: making the spatial correlation
isotropic (the default) instead of anisotropic and structuring the
spatiotemporal random fields as AR(1).

```{r dog-ar1, results='hide', echo=TRUE}
fit_dpl_iso <- update(fit_dpl, anisotropy = FALSE)
fit_dpl_ar1 <- update(fit_dpl, spatiotemporal = list("off", "ar1"))
```

```{r dog-aci2, echo=TRUE}
AIC(fit_dpl_ar1, fit_dpl, fit_dpl_iso)
```

We find that the anisotropic AR(1) is favoured. On inspecting the model output
with `summary()`, we notice that the spatial random field SD has become very
small once AR(1) spatiotemporal fields were added. We therefore consider
a model where these are turned off in the second component of the delta model:

```{r dog-ar1-2, echo=TRUE, results="hide"}
fit_dpl_ar1_only <- update(fit_dpl_ar1, spatial = list("on", "off"))
```

```{r dog-aic3, echo=TRUE}
AIC(fit_dpl_ar1_only, fit_dpl_ar1, fit_dpl)
```

```{r dog-print, echo=TRUE}
fit <- fit_dpl_ar1_only
sanity(fit)
summary(fit)
```

```{r dog-aniso, echo=TRUE}
plot_anisotropy(fit_dpl)
```

```{r dog-grid, echo=TRUE}
head(wcvi_grid, n = 3)
grid <- replicate_df(wcvi_grid, "year", time_values = unique(dat$year))
grid$log_depth <- log(grid$depth)
head(grid, n = 3)
```

```{r dog-pred1, echo=TRUE}
pred <- predict(fit, newdata = grid, type = "response")
```

```{r dog-pred2, echo=TRUE}
names(pred)
```

```{r plot-map, echo=TRUE}
plot_map <- function(dat, column) {
  ggplot(dat, aes(X, Y, fill = {{ column }})) +
    geom_raster() +
    facet_wrap(vars(year)) +
    coord_fixed()
}
```

```{r dog-plot1, echo=TRUE}
pred |> filter(year %in% c(2004, 2014, 2022)) |> 
  plot_map(est) +
  labs(fill = "Density") +
  scale_fill_viridis_c(trans = "log10")
```

```{r dog-plot2, echo=TRUE}
pred |> filter(year %in% 2004) |> 
  plot_map(omega_s2) +
  scale_fill_gradient2()
```

```{r dog-plot3, echo=TRUE}
pred |> filter(year %in% c(2004, 2022)) |> 
  plot_map(epsilon_st2) +
  scale_fill_gradient2()
```

```{r dog-plot4, echo=TRUE}
pred |> filter(year %in% 2004) |> 
  plot_map(est_non_rf1) +
  scale_fill_viridis_c(trans = "log10")
```

```{r dog-depth, echo=TRUE}
nd <- data.frame(
  log_depth = seq(min(dat$log_depth), max(dat$log_depth), length.out = 100), 
  year = max(dat$year)
)
pred_depth <- predict(fit, newdata = nd, se_fit = TRUE, re_form = NA, model = NA)
ggplot(pred_depth, aes(
  exp(log_depth), exp(est), 
  ymin = exp(est - 2 * est_se), 
  ymax = exp(est + 2 * est_se))) +
  geom_ribbon(fill = "grey90") +
  geom_line() +
  labs(x = "Depth (m)", y = "Density")
```

```{r dog-index, echo=TRUE}
pred2 <- predict(fit, newdata = grid, return_tmb_object = TRUE)
ind <- get_index(pred2, bias_correct = TRUE, area = rep(4, nrow(grid)))
```

```{r dog-index-plot, echo=TRUE}
ggplot(ind, aes(year, est, ymin = lwr, ymax = upr)) +
  geom_pointrange() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.03)), limits = c(0, NA)) +
  labs(y = "Biomass", x = "Year")
```

# Spatially varying coefficient example with citizen science data

Snowy Owls (*Bubo scandiacus*) breed on the arctic tundra and are irruptive migrants, meaning that they appear across the mid-latitudes of North America in much greater numbers in some winters than others.
The reasons for this interannual variation in the number of individuals migrating south are not well understood but seem to be related to high abundances of food during the breeding season and therefore sharp increases in breeding ground population densities [@robillard2016].
The North Atlantic Oscillation Index (NAO) has been linked to productivity of both owls and their prey in Europe [@millon2014].
Because both productivity and the choice of wintering location could be influenced by climate, we modelled a spatially varying effect of annual mean NAO index on winter abundance across the southern boundary of their winter distribution.
We fit counts observed in North America during annual Christmas Bird Counts [@cbc] using a negative binomial [NB2, @hilbe2011] distribution, random intercepts for year, spatial and spatiotemporal random fields, and a spatially varying coefficient associated with the NAO.

We can write the model as TODO

```{r owl-data}
snow <- readRDS(here::here("data/SNOW_data.rds"))
```

```{r owl-fit, echo=TRUE, results = "hide"}
mesh_snow <- make_mesh(snow, xy_cols = c("X", "Y"), cutoff = 1.5)
fit_owl <- sdmTMB(
  count ~ 1 + nao + (1 | year_f),
  spatial_varying = ~ nao,
  time = "year",
  data = snow, 
  family = nbinom2(link = "log"),
  spatial = "on",
  spatiotemporal = "iid",
  mesh = mesh_snow,
  reml = TRUE,
  silent = FALSE
)
```

```{r owl-sanity, eval=FALSE}
sanity(fit_owl)
```

```{r owl-print, echo=TRUE}
summary(fit_owl)
```

We can extract the SVC effect with TODO

```{r p-mean, message=FALSE, echo=TRUE, eval=TRUE, cache=TRUE}
nsim <- 200
zeta_s <- predict(fit_owl, newdata = snow, nsim = nsim, sims_var = "zeta_s")
sims <- spread_sims(fit_owl, nsim = nsim)
combined <- sims$nao + t(zeta_s)
snow$nao_effect <- apply(exp(combined), 2, median)
```

We can make a basic plot using the following code. A more elaborate version is shown in Figure \@ref(fig:owl-plot-fancy).

```{r owl-plot-basic, echo=TRUE, eval=FALSE}
ggplot(snow, aes(X, Y)) +
  geom_point(aes(colour = nao_effect)) +
  scale_colour_viridis_c()
```

Overall, we found a weak average positive effect of annual mean NAO on overall counts, but a southeast to northwest gradient in the intensity of the effect (Fig.\ \ref{fig:owl-nao}).
This result is consistent with owls closest to the Atlantic coast and those migrating the furthest south being the most affected by NAO.
We provide an extended version of this example in Appendix \ref{app:owls}.

```{r shapes, echo=FALSE}
if (!file.exists(here::here("data/ne_10m_lakes"))) {
  zip_file <- paste0("https://www.naturalearthdata.com/http//www.naturalearthdata.com/",
    "download/10m/physical/ne_10m_lakes.zip")
  download.file(zip_file, destfile = here::here("data/ne_10m_lakes.zip"))
  unzip(here::here("data/ne_10m_lakes.zip"), exdir = here::here("data/ne_10m_lakes"))
}
```

```{r shapes-read, echo=FALSE}
Albers <- "+proj=aea +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"
coast <- rnaturalearth::ne_coastline(scale = "medium", returnclass = "sf") %>%
  sf::st_transform(crs = Albers)
lakes <- sf::st_read(here::here("data/ne_10m_lakes"), quiet = TRUE)
lakes <- lakes[lakes$scalerank == 0, ] %>% sf::st_transform(crs = Albers)
```

```{r owl-proj2}
# project to get nice axis limits 
snow2 <- snow |> mutate(X = X * 100000, Y = Y * 100000)
snow2 <- snow2 |> mutate(x = X, y = Y) |> 
  sf::st_as_sf(coords = c("x", "y"), crs = Albers)
```

```{r owl-plot-fancy, echo=FALSE, eval=TRUE, fig.cap="Map of spatially varying coefficient effects of NAO on Snowy Owl count. In the west, the lower values are near 1 implying no effect, whereas in the southeast the effect becomes positive. Points represent all count locations."}
nao_effect_df <- select(snow, X, Y, nao_effect) |> distinct() |> 
  mutate(X = X * 100000, Y = Y * 100000) # to match CRS
ggplot(data = nao_effect_df) +
  geom_point(aes(X, Y, colour = nao_effect)) +
  geom_sf(data = coast, colour = "gray50") +
  geom_sf(data = lakes, colour = "gray50", fill = NA) +
  coord_sf(
    xlim = c(min(snow2$X), max(snow2$X)),
    ylim = c(min(snow2$Y), max(snow2$Y))
  ) +
   scale_colour_viridis_c(
    guide = guide_colourbar(direction = "horizontal", title.position = "left")
  ) +
  labs(colour = "Estimated multiplicative effect\nof NAO on Snowy Owl count") +
    theme(legend.position = "bottom", axis.title = element_blank())
```

\clearpage

\pkg{sdmTMB} can be installed from the Comprehensive \proglang{R} Archive Network (CRAN) at <https://CRAN.R-project.org/package=sdmTMB>

```{r sdmTMB-install, eval=FALSE, echo=TRUE}
install.packages("sdmTMB", dependencies = TRUE)
```

Importantly, `dependencies = TRUE` includes the \proglang{R}-INLA package  [@rue2009; @lindgren2011; @lindgren2015], which is not on CRAN but is needed for the `make_mesh()` function below.

Here, we illustrate a spatial model fit to Pacific Cod (*Gadus macrocephalus*) trawl survey data from Queen Charlotte Sound, British Columbia, Canada.
Our model contains a main effect of depth as a penalized smoother, a spatial random field, and Tweedie observation error.
Our data frame `pcod` (built into the package) has a column `year` for the year of the survey, `density` for biomass density of Pacific Cod in the area swept for a given survey tow, `depth` for depth in meters of that tow, and spatial longitude and latitude `lon` and `lat`.

```{r sdmTMB-lib0, eval=FALSE, echo=FALSE}
# library("sdmTMB")
# # simplify for paper:
pcod_original <- pcod # save it for later
# pcod <- dplyr::select(pcod, year, density, depth, lon, lat) |>
#   tibble::as_tibble()

```

```{r sdmTMB-lib, eval=FALSE, echo=TRUE}
library("sdmTMB")
dplyr::select(pcod, year, density, depth, lon, lat) |>
  head(n = 3)
```

<!-- An \pkg{sdmTMB} model requires a data frame that contains a response column, columns for any predictors, and columns for spatial coordinates. -->
Usually it makes sense to convert the spatial coordinates to an equidistant projection such as the Universal Transverse Mercator (UTM) to ensure that distance remains constant throughout the study region [e.g., using `sf::st_transform()`, @pebesma2018].
Here we use the helper function `add_utm_columns()` to add UTM coordinates with km units (so our estimated spatial range parameter is not too big or small).
By default, the function will guess our UTM zone and create new columns `X` and `Y`.

```{r pcod-utms-eval, echo=TRUE, eval=FALSE}
pcod <- add_utm_columns(pcod, c("lon", "lat"), units = "km")
```

```
#> Detected UTM zone 9N; CRS = 32609.
#> Visit https://epsg.io/32609 to verify.
```

```{r pcod-head2, echo=TRUE, eval=FALSE}
head(pcod, n = 3)
```

We then create a mesh object that contains triangulation and projection matrices needed to apply the SPDE approach.

```{r, echo=TRUE, eval=FALSE, cache=TRUE}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)
```

Here, `cutoff` defines the minimum allowed distance between mesh vertices in the units of `X` and `Y`.
`make_mesh()` is a wrapper function for `INLA::inla.mesh.create()` and `cutoff` is passed through to it.
Alternatively, we could have created any mesh via the \proglang{R}-\pkg{INLA} package and supplied it to the `mesh` argument in `make_mesh()`.
We can inspect our mesh object with the associated plotting method (`plot(mesh)`; see Fig. \ref{fig:pcod-fig}A).

We can then fit a model with spatial random fields (`spatial = "on"`) via the function `sdmTMB()`.
We use a penalized smoother for depth as a main effect via `s()` from the \pkg{mgcv} package.
We specify the family as Tweedie [@tweedie1984] to account for positive continuous density values that also contain zeros.
An alternative would be the `delta_gamma()` family [@aitchison1955] or to model catch weight with an offset for log(area swept) (via the `offset` argument).

```{r pcod-eg1-fit, warning=FALSE, message=FALSE, echo=TRUE, eval=FALSE, cache=TRUE}
fit <- sdmTMB(
  density ~ s(depth),
  data = pcod,
  family = tweedie(link = "log"),
  mesh = mesh,
  spatial = "on"
)
```

We can get a summary of the fit with the `print()` method:

```{r pcod-eg1-summary,eval=FALSE, echo=TRUE}
fit
```

The output indicates our model was fit by maximum (marginal) likelihood (`ML`).
We also see the formula, mesh, fitted data, and family.
Next we see any estimated main effects including the linear component of the smoother (`sdepth`), the standard deviation on the smoother weights (`sds(depth)`), the Tweedie dispersion and power parameters, the Matérn range distance, the marginal spatial field standard deviation, and the negative log likelihood at convergence.
The smoother summary follows the format used in the \pkg{brms} package [@burkner2017].
As noted in the print method, we can extract a tidy data frame similar to the format in the \pkg{broom} [@broom] and \pkg{broom.mixed} [@broommixed] packages:

```{r pcod-eg1-tidy-fe, eval=FALSE, echo=TRUE}
tidy(fit, conf.int = TRUE)
```

```{r pcod-eg1-tidy-re, eval=FALSE, echo=TRUE}
tidy(fit, effects = "ran_pars", conf.int = TRUE)
```

The latter `tidy()` call returns all variance components of the model and omits the standard errors since these are calculated in log space.

We can run some basic checks on our model with the `sanity()` function:

```{r pcod-eg1-sanity, eval=FALSE, echo=TRUE}
sanity(fit)
```

```
#> ✓ Non-linear minimizer suggests successful convergence
#> ✓ Hessian matrix is positive definite
#> ✓ No extreme or very small eigenvalues detected
#> ✓ No gradients with respect to fixed effects are >= 0.001
#> ✓ No fixed-effect standard errors are NA
#> ✓ No fixed-effect standard errors look unreasonably large
#> ✓ No sigma parameters are < 0.01
#> ✓ No sigma parameters are > 100
#> ✓ Range parameter doesn't look unreasonably large
```

We can make predictions with the `predict()` method (`?predict.sdmTMB`) and optionally use the `newdata` argument to predict on a new data frame with any locations and values for the predictor columns.
Here, we will predict on a 2 $\times$ 2 km grid (`qcs_grid`) that covers the entire region of interest so we can visualize the predictions spatially.
The grid contains spatial covariate columns and all predictors used in the model set at values for which we want to predict.

```{r pcod-eg1-suppress-depth2, eval=FALSE, echo=FALSE}
qcs_grid_original <- qcs_grid
qcs_grid$depth_scaled <- NULL # for simplicity in paper
qcs_grid$depth_scaled2 <- NULL # for simplicity in paper
```

```{r qcs-head, echo=FALSE}
head(qcs_grid, n = 3)
```

```{r pcod-eg1-predict, echo=FALSE, eval=TRUE, cache=TRUE}
p <- predict(fit, newdata = qcs_grid)
```

The output of `predict()` is a data frame containing overall estimates in link space (`est`), estimates from the non-random-field components (`est_non_rf`; here, intercept and depth), estimates from the random field components combined (`est_rf`), and estimates from the individual random field components (here, `omega_s`---the spatial field).

```{r pcod-p-tibble, eval=FALSE}
p <- tibble::as_tibble(p)
```

```{r pcod-head-predict, echo=TRUE, eval=FALSE}
head(p, n = 3)
```

We show a basic plot of the estimated spatial random field, predictions across a depth gradient, and predictions in space in Fig.\ \ref{fig:pcod-fig}B--D.

\begin{figure}
\centering
\includegraphics[width=5.2in]{../../figs/pcod-fig}
\caption[]{Output from the Pacific Cod spatial model example in Queen Charlotte Sound, BC, Canada.
(A) SPDE mesh (lines) combined with the trawl survey observations (points). Finer meshes will be slower to fit but generally increase the accuracy of the SPDE approximation. The circle area corresponds to biomass density of Pacific Cod caught on individual trawls.
(B) Spatial random field: these values are shown in link (log) space and represent spatially correlated deviations that are not accounted for by the depth effect. (C) Overall prediction: these estimates represent the combination of all fixed and random effects. Crosses illustrate center of gravity (\texttt{get\_cog()}) from a spatiotemporal version; lighter crosses are more recent.
(D) Conditional effect of depth modelled as a penalized smoother.
These predictions are made omitting the spatial random field.
(E) Standardized area-weighted population index derived from a spatiotemporal version (\texttt{get\_index()}). The line represents the estimate and the ribbon indicates a 95\% confidence interval ($\pm$ 2 SEs).}\label{fig:pcod-fig}
\end{figure}

We could extend this to a spatiotemporal model simply by supplying the year column name to the `time` argument (`time = "year"`) and specifying how we want the spatiotemporal random fields to be structured.
Here, we will keep the spatial field (`spatial = "on"`), structure the spatiotemporal random fields as independent (`spatiotemporal = "iid"`), include a time-varying intercept (`~ 1`) fit as a random walk process (`time_varying_type = "rw"`), include anisotropy to allow spatial correlation to decay at different rates up the coast vs.\ away from the coast (`anisotropy = TRUE`), and use `silent = FALSE` to see model fitting progress.

\clearpage

```{r time1, cache=TRUE, eval=FALSE}
t1 <- Sys.time()
```

```{r fit-tv, eval=FALSE, echo=TRUE, cache=TRUE, results='hide'}
fit_spatiotemporal <- sdmTMB(
  density ~ 0 + s(depth, k = 5),
  time_varying = ~ 1,
  time_varying_type = "rw",
  family = tweedie(link = "log"),
  data = pcod,
  mesh = mesh,
  time = "year",
  spatial = "on",
  spatiotemporal = "iid",
  anisotropy = TRUE,
  silent = FALSE
)
```

```{r time2, cache=TRUE, eval=FALSE}
t2 <- Sys.time()
```

```{r fit-tv-print, eval=FALSE, echo=TRUE}
fit_spatiotemporal
```

We could inspect the anisotropy with `plot_anisotropy()`, which plots an ellipse showing the spatial range with the spatial coordinates centered on zero.
We could generate an area-weighted population index (e.g., a relative or absolute index of abundance or biomass) that is independent of sampling locations by predicting from that fit on a grid covering the area of interest and summing the predicted biomass with the `get_index()` function (Fig.\ \ref{fig:pcod-fig}E, Appendix \ref{app:pcod}).

```{r pcod-restore, echo=FALSE, eval=FALSE}
pcod <- pcod_original
qcs_grid <- qcs_grid_original
```

# Spatially varying coefficient example with citizen science data

Snowy Owls (*Bubo scandiacus*) breed on the arctic tundra and are irruptive migrants, meaning that they appear across the mid-latitudes of North America in much greater numbers in some winters than others.
The reasons for this interannual variation in the number of individuals migrating south are not well understood but seem to be related to high abundances of food during the breeding season and therefore sharp increases in breeding ground population densities [@robillard2016].
The North Atlantic Oscillation Index (NAO) has been linked to productivity of both owls and their prey in Europe [@millon2014].
Because both productivity and the choice of wintering location could be influenced by climate, we modelled a spatially varying effect of annual mean NAO index on winter abundance across the southern boundary of their winter distribution.
We fit counts observed in North America during annual Christmas Bird Counts [@cbc] using a negative binomial [NB2, @hilbe2011] distribution, random intercepts for year, spatial and spatiotemporal random fields, and a spatially varying coefficient associated with the NAO.

```{r owl-fit-old, warning=FALSE, message=FALSE, cache=TRUE, echo=TRUE, eval=FALSE}
mesh <- make_mesh(snow, xy_cols = c("X", "Y"), cutoff = 1.5)
fit_owls <- sdmTMB(
  count ~ nao + (1 | year_factor),
  spatial_varying = ~ nao,
  family = nbinom2(link = "log"),
  data = snow,
  mesh = mesh,
  time = "year",
  spatial = "on",
  spatiotemporal = "iid"
)
```

We found a weak average positive effect of annual mean NAO on overall counts, but a southeast to northwest gradient in the intensity of the effect (Fig.\ \ref{fig:owl-nao}).
This result is consistent with owls closest to the Atlantic coast and those migrating the furthest south being the most affected by NAO.
We provide an extended version of this example in Appendix \ref{app:owls}.

\clearpage

```{r owl-nao, fig.cap="Spatially varying coefficient for effect of mean annual NAO (North Atlantic Oscillation) on counts of Snowy Owls observed on annual Christmas Bird Counts 1979--2020 in Canada and the US. Points represent all count locations and circle area is scaled to the mean number of owls observed per year (range: 0 to 8). The effect is multiplicative on owl count per NAO unit.", out.width="4.1in", fig.align='center'}
owl <- here::here("figs", "owl-nao-effect.png")
knitr::include_graphics(owl)
```

# Model validation and selection

Validation and selection of state-space models is challenging, particularly when using the Laplace approximation [@thygesen2017a].
We provide several approaches to assist this process:
(1) The Akaike Information Criterion [AIC, @akaike1974] can be calculated with `AIC()`, although AIC has well-documented biases with mixed-effects models [@liang2008].
(2) Alternatively, k-fold cross validation with `sdmTMB_cv()` can be used with user-specified or randomly chosen folds for model selection [e.g., via expected log predictive density [ELPD], @vehtari2017] or to evaluate goodness of fit according to user-calculated criteria (e.g., mean squared error, area under the curve).
(3) An \pkg{sdmTMB} model can be passed to the \pkg{tmbstan} package [@monnahan2018] to sample from the joint posterior with Stan [@carpenter2017], evaluate the accuracy of the Laplace approximation, or perform posterior predictive checks (see `?extract_mcmc`).
(4) The `residuals()` method by default returns randomized quantile [@dunn1996] or probability integral transform (PIT) [@smith1985] residuals.
<!-- These residuals, however, are calculated as $\hat{y} - y$ where $\hat{y}$ is the sum of the fixed effect estimates and the random effect values that maximize the likelihood conditional on the estimated fixed effects.  -->
For state-space models, these residuals have known statistical issues with the Laplace approximation [@thygesen2017a] but are quick to calculate.
A version that uses Markov chain Monte Carlo (MCMC) sampling of the random effects to avoid this issue [@rufener2021] is recommended but slower (`?residuals.sdmTMB`).
Simulation-based residuals [@dharma] are also possible.
(5) The `simulate.sdmTMB()` method can simulate from fitted models and the `sdmTMB_simulate()` function facilitates simulating data without starting from a fitted model.
Models can be fit to these simulated data to ensure identifiability, evaluate bias and precision in parameter estimation, or evaluate the consequences of model misspecification.

\clearpage

# Package comparisons

There are many \proglang{R} packages capable of fitting geostatistical spatial or spatiotemporal models [e.g., @heaton2019].
\pkg{sdmTMB}, \pkg{VAST}, \proglang{R}-\pkg{INLA}/\pkg{inlabru}, and \pkg{spaMM} [@rousset2014] are the most closely related, as they all provide a user interface to SPDE-based GRF models.
In our software comparison (Table \ref{tab:functionality}), we also include \pkg{mgcv} as it can be adapted to use the SPDE [@miller2019] and \pkg{spBayes} [@finley2007; @finley2015] since it is a prominent package that can fit related predictive-process models without the SPDE.
\pkg{sdmTMB}, \pkg{VAST}, and \pkg{mgcv} can estimate anisotropic covariance whereas \proglang{R}-\pkg{INLA}/\pkg{inlabru} and \pkg{spBayes} are currently limited to isotropic covariance.
\pkg{sdmTMB} and \pkg{mgcv} focus on univariate response data, whereas \pkg{VAST}, \proglang{R}-\pkg{INLA}/\pkg{inlabru}, \pkg{spaMM}, and \pkg{spBayes} extend to multivariate responses with various limitations.
To our knowledge, \pkg{VAST} is the only package to implement spatial [@thorson2015] and spatial dynamic factor analysis [@thorson2016] and spatial empirical orthogonal function (EOF) regression [@thorson2020].
Of these packages, only \pkg{sdmTMB} and \pkg{inlabru} can currently fit threshold (e.g., hockey-stick) covariate relationships.
\pkg{spaMM} is limited to a spatial random field (i.e., does not fit spatiotemporal fields) and \pkg{spBayes} implements spatiotemporal fields, but only as a random walk.
There is considerable variability in the available observation likelihoods across packages (Table \ref{tab:functionality}).
We provide comparisons of the syntax and the reproducibility of results from models fit using \pkg{sdmTMB}, \proglang{R}-\pkg{INLA}, and \pkg{inlabru} in Appendix&nbsp;\ref{app:inla}.
<!-- and \pkg{VAST} (Appendix 5). -->

```{r compare-table, results='asis', eval=FALSE}
library(kableExtra)
suppressMessages(suppressWarnings(d <- readr::read_csv(here::here("doc/software-comparison.csv"), show_col_types = FALSE, comment = "#")))
d <- as.matrix(d)
# d[d == "1"] <- "\\checkmark"
d <- gsub("^1", "\\\\checkmark", d)
# d[d == "1*"] <- "\\checkmark\\footnote{a}"
# d[d == "GAMs*"] <- "GAMs\\footnote{a}"
# d[d == "0"] <- "--"
d <- gsub("^0", "--", d)
# d[d == "0*"] <- "--\\footnote{a}"
d <- as.data.frame(d)
names(d)[1] <- ""
row.names(d) <- NULL
d <- d[,c(1, 2, 3, 5, 6, 4, 7, 8, 9)]
d$Hmsc <- NULL

d$INLA <- NULL
names(d)[names(d) == "inlabru"] <- "INLA/inlabru"

knitr::kable(d, format = "latex", caption = "Comparison of functionality between several R packages that can fit geostatistical GLMMs.",
  booktabs = TRUE,
  linesep = c(
    rep('', 4), # coefs
    '\\addlinespace',
    rep('', 6), # correlation stuff
    '\\addlinespace',
    rep('', 12), # likelihood
    '\\addlinespace',
    rep('', 2), # bayes
    '\\addlinespace'
  ),
  escape = FALSE, row.names = FALSE) %>%
  kableExtra::kable_styling(font_size = 8) %>%
  kableExtra::footnote(
    general = paste(
      "$^1$Technically possible but non-trivial.",
      "$^2$Penalized smoother GAMs that determine `wiggliness'.",
      "$^3$inlabru but not INLA.",
      "$^4$Spatiotemporal fields as random walk only.",
      "$^5$SPDE approach as in Miller et al. (2019).",
      "$^6$Does have predictive process knots.",
      "$^7$Zero-inflated NB2 only.",
      "$^8$Tweedie power parameter fixed for \\\\texttt{mgcv::gamm()}.",
      "$^9$Possible as log-linked Poisson GLMM with aggregated data.",
      "$^{10}$Hurdle models possible by fitting components separately."
    ),
    escape = FALSE, threeparttable = TRUE
  )
```

\begin{table}

\centering
\fontsize{8}{10}\selectfont
\begin{threeparttable}
\begin{tabular}[t]{lllllll}
\toprule
 & \pkg{sdmTMB} & \pkg{VAST} & \proglang{R}-\pkg{INLA}/\pkg{inlabru} & \pkg{mgcv} & \pkg{spBayes} & \pkg{spaMM}\\
\midrule
Time-varying coefficients & \checkmark & --$^1$ & \checkmark & \checkmark & \checkmark & --\\
Spatially varying coefficients (SVC) & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & --\\
GAMs$^2$ & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Threshold covariates & \checkmark & -- & \checkmark$^3$ & -- & -- & --\\
Offsets & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark\\
\addlinespace
Spatiotemporal fields & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark$^4$ & --\\
Spatial + spatiotemporal fields & \checkmark & \checkmark & \checkmark & \checkmark & -- & --\\
Anisotropy & \checkmark & \checkmark & -- & \checkmark & -- & --\\
Correlation barriers & \checkmark & \checkmark & \checkmark & \checkmark & -- & --\\
Separate range parameters for fields & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Share range parameters across fields & \checkmark & \checkmark & \checkmark & -- & -- & --\\
SPDE-based & \checkmark & \checkmark & \checkmark & \checkmark$^5$ & --$^6$ & \checkmark\\
\addlinespace
NB1 distribution & \checkmark & -- & \checkmark & \checkmark & -- & \checkmark\\
NB2 distribution & \checkmark & \checkmark$^7$ & \checkmark & \checkmark & -- & \checkmark\\
Zero-truncated distributions & \checkmark & -- & \checkmark & -- & -- & \checkmark\\
Zero-inflated distributions & \checkmark & \checkmark & \checkmark & -- & -- & \checkmark\\
Tweedie distribution & \checkmark & \checkmark & \checkmark & \checkmark$^8$ & -- & --\\
Student-t distribution & \checkmark & -- & \checkmark & \checkmark & -- & --\\
Censored Poisson distribution & \checkmark & -- & \checkmark & -- & -- & --\\
log Gaussian Cox processes & --$^9$ & --$^9$ & \checkmark & --$^9$ & --$^9$ & --$^9$\\
Multivariate responses & -- & \checkmark & \checkmark & -- & \checkmark & \checkmark\\
Built-in delta/hurdle models & \checkmark & \checkmark & \checkmark & --$^{10}$ & -- & \checkmark\\
Poisson-link delta model & \checkmark & \checkmark & \checkmark & -- & -- & --\\
Likelihood weights & \checkmark & -- & \checkmark & \checkmark & \checkmark & \checkmark\\
Maximum/marginal likelihood & \checkmark & \checkmark & -- & \checkmark & -- & --\\
\addlinespace
Bayesian/optionally Bayesian & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & --\\
Priors/penalties & \checkmark & -- & \checkmark & -- & \checkmark & --\\
Matern PC priors & \checkmark & -- & \checkmark & -- & -- & --\\
\addlinespace
Spatial (or spatial dynamic) factor analysis & -- & \checkmark & -- & -- & -- & --\\
Empirical Orthogonal Function (EOF) analysis & -- & \checkmark & -- & -- & -- & --\\
Built-in area-weighted index standardization & \checkmark & \checkmark & -- & -- & -- & --\\
Built-in cross-validation & \checkmark & -- & -- & -- & -- & --\\
\bottomrule
\end{tabular}
\end{threeparttable}
\label{tab:functionality}

\caption{Comparison of functionality between several \proglang{R} packages that can fit geostatistical GLMMs.
Notes:
$^1$Technically possible but non-trivial.
$^2$Penalized smoother GAMs that determine `wiggliness'.
$^3$\pkg{inlabru} but not \proglang{R}-\pkg{INLA}.
$^4$Spatiotemporal fields as random walk only.
$^5$SPDE approach as in \citet{miller2019}.
$^6$Does have predictive process knots.
$^7$Zero-inflated NB2 only.
$^8$Tweedie power parameter fixed for \texttt{mgcv::gamm()}.
$^9$Possible as log-linked Poisson GLMM with aggregated data.
$^{10}$Hurdle models possible by fitting components separately.
}
\end{table}

<!-- TODO: Line 242. Can \pkg{sdmTMB} be run in parallel? If so, does the parallelism share memory across the cores or does the memory grow with the number of cores? This is a HUGE feature of PARDISO within \proglang{R}-\pkg{INLA}/\pkg{inlabru} allowing for very large models to be fit on HPCs. It might be worth mentioning and adding to Table 1. Also Line 242: do you expect the same relative performance between the software packages as more cores are used? -->

We ran a simple speed comparison between \pkg{sdmTMB}, \proglang{R}-\pkg{INLA}/\pkg{inlabru}, and \pkg{mgcv} for fitting an SPDE spatial random field model to 1000 data points with Gaussian error across a range of mesh resolutions (Fig. \ref{fig:timing}, Appendix \ref{app:speed}).
In this test, \pkg{sdmTMB} was the fastest across all mesh resolutions; however, \proglang{R}-\pkg{INLA}/\pkg{inlabru} and \pkg{spaMM} were less affected by mesh resolution than \pkg{sdmTMB}.
\pkg{mgcv} was most affected by mesh resolution.
Our test was restricted to one core and default \proglang{R} algebra libraries; all packages could run faster with optimized libraries and parallel processing.
Results with optimized math libraries on one core (\pkg{openBLAS}: @openblas and \pkg{PARDISO}: @pardiso7.2a) resulted in a ~10% speed increase for \pkg{sdmTMB} and \pkg{inlabru} and a ~7--9-fold speed increase for \pkg{mgcv}.

\begin{figure}[htb]
\centering
\includegraphics[width=3.5in]{../../figs/timing-spatial-2023-03-16-xkcd.pdf}
\caption[]{Comparison of time to fit an SPDE spatial random field model with 1000 observations, an intercept and one predictor, Gaussian error, and a sequence of SPDE resolutions. Lines represent means and ribbons 95\% quantiles across 50 random iterations. Note the log-distributed x axis. \pkg{VAST} should be similar to \pkg{sdmTMB} and so is not shown. \pkg{inlabru} used the empirical Bayes integration strategy and Gaussian approximation with \texttt{bru\_max\_iter = 1}, and the \texttt{like()} formulation. \pkg{mgcv} used \texttt{bam()}, \texttt{method = `fREML'}, and discretized covariates \citep{miller2019}. Note that \pkg{spaMM} only fits spatial, not spatiotemporal, models. All platforms were restricted to one core and could be faster with parallel computation or optimized algebra libraries.}\label{fig:timing}
\end{figure}

# Discussion

How does one choose among the related packages mentioned in this paper to fit SPDE-based geostatistical GLMMs?
Assuming a given package can fit the model of interest (Table  &nbsp;\ref{tab:functionality}), we suggest the primary differences are the user interface (Appendix \ref{app:inla}) and speed.
We think users familiar with `stats::glm()`, \pkg{lme4}, \pkg{mgcv}, or \pkg{glmmTMB} will find \pkg{sdmTMB} most approachable.
Users familiar with \proglang{R}-\pkg{INLA} will find \pkg{inlabru} approachable.
Users familiar with \pkg{mgcv} can adapt \pkg{mgcv} to fit similar models with custom code [@miller2019] and \proglang{R}-\pkg{INLA}/\pkg{inlabru} and \pkg{mgcv} are also general purpose modelling packages.
\pkg{VAST} is the sole option for fitting some multivariate models;
alternatively, because \pkg{VAST} focuses on multivariate delta models and fisheries applications, users fitting "simple" univariate spatial/spatiotemporal GLMMs in non-fisheries contexts may find \pkg{sdmTMB} more straightforward.
Users looking for calculation, with uncertainty, of derived variables such as area-weighted population indexes, may favour \pkg{sdmTMB} or \pkg{VAST} (although such quantities can be post hoc derived with other packages).
<!-- Speed-wise, the TMB-based packages \pkg{sdmTMB} and \pkg{VAST} were fastest up to at least 1000 mesh nodes, which in practice can cover fairly complex spatial processes on large datasets [e.g., @anderson2019b; @maureaud2021; @english2022]. -->

With respect to computational speed, \pkg{sdmTMB} (and due to having the same \pkg{TMB} backend, \pkg{VAST}) were fastest with a spatial model up to at least \textasciitilde 1000 mesh vertices at approximately a seven-times (at 1165 vertices) to 30-times (at 125 vertices) speed increase over \proglang{R}-\pkg{INLA}/\pkg{inlabru}.
These speed increases can allow for more rapid and thorough model exploration and experimentation with a class of computationally intensive models.
However, for users ultimately interested in Bayesian inference, the approximate Bayesian inference offered by \proglang{R}-\pkg{INLA}/\pkg{inlabru} is likely to be considerably faster than passing the same model from \pkg{sdmTMB}/\pkg{VAST} to \pkg{tmbstan} for full Bayesian inference.
<!-- Furthermore, although both \proglang{R}-\pkg{INLA}/\pkg{inlabru} and \pkg{sdmTMB} can use parallel processing, the PARDISO library [@pardiso-7.2c] within \proglang{R}-\pkg{INLA}/\pkg{inlabru} allows memory to be shared across cores. -->

<!-- However, with the SPDE, \pkg{mgcv} is typically slower than \pkg{sdmTMB} for similar models without optimized matrix libraries and lacks certain  -->

<!--
There are many \proglang{R} packages available for fitting spatially explicit models to large datasets and we cannot review them all here [see @heaton2019]; however, the three most closely related are \pkg{VAST}, \pkg{inlabru}, and \pkg{mgcv} (Table 1).
\pkg{VAST} is also built on \pkg{TMB} and, for univariate data, should be able to fit equivalent models with some minor differences in the user interfaces.
Interface-wise, \pkg{VAST} is less modular (e.g., by default constructing meshes before fitting, making predictions at the same time as fitting), is written from a fisheries perspective, and lacks some interface elements that we think make \pkg{sdmTMB} more broadly usable by applied ecologists.
For example, `sdmTMB()` has a `family` argument as in `glm()`, a `predict()` method, and human-readable options for spatial/spatiotemporal fields (e.g., `spatial = 'on'`, `spatiotemporal = 'AR1'`).
From a fisheries perspective, \pkg{sdmTMB} achieves three of the four design principles outlined in @thorson2019 for index standardization: it allows area weighting, can distinguish catchability and habitat covariates (by deciding what is fixed when predicting), and conditions on missing covariates (random fields).
Unlike \pkg{VAST}, however, \pkg{sdmTMB} does not currently extend to multivariate applications or support delta/hurdle models.

\pkg{inlabru} is similar to \pkg{sdmTMB} in design, but differs fundamentally in that fitting is achieved via \proglang{R}-\pkg{INLA} and approximate Bayesian computation.
For some models, TMB/\pkg{sdmTMB} can be considerably faster than \proglang{R}-\pkg{INLA}/\pkg{inlabru}, but those differences depend on the SPDE mesh, model, algebra libraries, and fitting algorithm settings [Fig. \@ref(fig:timing), @osgood-zimmerman2021].
One major capability included in \pkg{sdmTMB} but not \proglang{R}-\pkg{INLA}/\pkg{inlabru} is anisotropy.
Furthermore, we think the \pkg{sdmTMB} interface will be more familiar than \pkg{inlabru} to users of `glm()`, \pkg{lme4}, \pkg{mgcv}, or \pkg{glmmTMB}.
On the other hand, \pkg{inlabru} fits additional model classes including log Gaussian Cox processes and may be particularly usable to those already familiar with \proglang{R}-\pkg{INLA}.

\pkg{mgcv} can fit many similar models to \pkg{sdmTMB} either with various flavours of splines or even with the SPDE approximation to random fields [@miller2019]; however, there are some notable advantages of \pkg{sdmTMB} for the specific class of models discussed here.
First, the smoothers typically used with \pkg{mgcv} (e.g., thin plate splines), while capable of making equivalent predictions [@miller2019; @pedersen2019], do not offer the ecological interpretability of a spatially explicit random field model (spatial range and variance estimates).
Second, the implementation of the Tweedie distribution in \pkg{mgcv} requires the power parameter to be fixed in a mixed effects model.
Third, \pkg{mgcv} does not have built-in functionality to calculate derived values with uncertainty such as population indexes or center of gravity.
Despite these \pkg{sdmTMB} advantages for the models discussed here, \pkg{mgcv} is an immensely powerful modelling package that extends well beyond spatial random field models.
-->

<!-- MVT? time-varying epsilon, non-gaussian fields -->
<!-- 3d models, continuous time, -->

Additional functionality in \pkg{sdmTMB} not already mentioned includes interpolating across missing time slices and forecasting, the barrier SPDE model [@bakka2019], time-varying spatiotemporal covariance parameters [@ward2022], and simulation from the parameter joint precision matrix.
Future development may include additional zero-inflated models, improvements to Bayesian sampling efficiency [e.g., @monnahan2021], continuous-time models [e.g., @blangiardo2015book], multivariate responses, and non-Gaussian random fields [e.g., @anderson2019].
The included \pkg{TMB} `.cpp` file provides a tested model template that can be modified to add additional features.

# Acknowledgements

\pkg{sdmTMB} would not be possible without the \pkg{TMB} [@kristensen2016] and \proglang{R}-\pkg{INLA} [@rue2009; @lindgren2011; @lindgren2015] \proglang{R} packages.
\pkg{sdmTMB} is heavily inspired by and in some places code has been adapted from both the \pkg{VAST} [@thorson2019] and \pkg{glmmTMB} [@brooks2017] \proglang{R} packages (as described in the DESCRIPTION and inst/COPYRIGHTS files).
Smoother support was possible thanks to \pkg{mgcv} [@wood2017a].
We thank the authors of all these packages.
We thank S. Kotwicki, M. Lindmark, M. Martin, C.C. Monnahan, P.M. Regular, J.T. Thorson, and J. Watson for helpful comments that substantially improved the manuscript.
Christmas Bird Count data were provided by National Audubon Society and through the generous efforts of Bird Studies Canada and countless volunteers across the Western Hemisphere.

\clearpage

\appendix

```{r child="pcod.Rmd", eval=FALSE, cache=TRUE}
```

```{r child="owls.Rmd", eval=FALSE, cache=TRUE}
```

```{r child="inla-comparison.Rmd", eval=FALSE, cache=TRUE}
```

```{r child="timing.Rmd", eval=FALSE, cache=TRUE}
```